### 2019-01-15 doc/papers/nn/manuscript.tex
Initial commit. Start writing papers.

\documentclass[a4paper]{article}

\usepackage{amsmath}
\usepackage{fullpage}

\title{TensorAlloy: a fast machine learning framework for alloys}

\author{Xin Chen}

\begin{document}

\maketitle

\section{Method}

\subsection{Atomic forces}

In the ANN model, the total energy, $E^{total}$, of a structure is the sum of 
atomic energies:

\begin{gather}
E^{total} = \sum_{i}^{N}{E_i} \\
E_i = \mathbf{NN}_\alpha (G_i)
\end{gather}

\noindent where $G_i$ is the atomic descriptor of $\alpha$-type atom $i$.

\subsection{Stress}

\end{document}