### 2019-06-26 doc/papers/nn/manuscript.tex
CPC major improvements, revision 1.

\documentclass[preprint]{revtex4-2}

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
\raggedbottom           % don't add extra vertical space
\begin{comment}
\pagestyle{empty}       % use if page numbers not wanted
\end{comment}

\bibliographystyle{apsrev4-2}

\begin{document}

\title{
    TensorAlloy: a highly efficient atomistic neural network program for alloys
}
\author{Xin Chen}
\author{Xing-Yu Gao}
\author{Ya-Fan Zhao}
\author{De-Ye Lin}
\email{lindeye0716@163.com}
\author{Wei-Dong Chu}
\author{Hai-Feng Song}
\email{song\_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Mathematics, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help. 
However, the traditional way to construct a 
machine learning model computation graph starts from pre-computed descriptors, 
but not positions, due to technical challenges. Thus, deriving analytical atomic 
force and virial stress of a machine learning model becomes quite difficult as 
the partial derivations of descriptors with respect to positions need separate 
implementations, which requires plenty of efforts and its efficiency can not be 
guaranteed.
In this paper, we designed a new method, named the virtual atom approach, for 
implementing the symmetry function descriptor and the corresponding atomistic 
neural networks. With the virtual atom approach, we are able to build 
computation graph from atomic positions \textemdash but not descriptors 
\textemdash to total energy directly, thus asking modern machine learning 
platforms to compute atomic force and virial stress becoming possible. We also 
derived a simple and machine learning friendly equation for calculating virial 
stress. All these new algorithms have been implemented in our Python package, 
TensorAlloy. TensorAlloy supports constructing machine learning interaction 
potentials for both \textit{molecules} and \textit{solids}.
\end{abstract}

\maketitle

\newcommand{\tensoralloy}{\textbf{TensorAlloy}}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 1. Introduction
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\newpage

\section{Introduction}

Atomistic simulations (\textit{ab initio} calculation, classic molecular 
dynamics simulation (MD), etc) are powerful tools for studying physical and 
chemical properties of materials. To obtain reliable simulation results, 
atomistic interactions must be properly described. The very accurate 
\textit{ab initio} and quantum chemical methods have gain significant 
improvements over the past decades but their applicabilities are restricted by 
computation expenses. Physical model based empirical potentials, such as the 
embedded-atom method\cite{eam-1,eam-2,eam-3,eam-4,eam-5} or its variants (
modified embedded-atom method\cite{eam-5,meam-1,meam-2,meam-3}, 
angular-dependent interaction potential\cite{adp-1,adp-2,adp-3}, etc), are much 
more popular for long-time simulation because of their reasonable accuraty and 
acceptable costs. 
However, designing empirical potentials and optimizing parameters for 
alloys are still challenging tasks\cite{eam-5, adp-2, SNAP_2018}.

In the past few years, machine learning (ML) has become one of the most 
influential topic in almost all fields of science. A lot of effort has been made 
by various physicists to explore the possiblity of using ML models \textemdash 
instead of traditional empirical potentials \textemdash to describe atomistic 
interactions during MD simulations\cite{kCON,SchNet,DTNN,SNAP_2018,SNAP_Mo_2017,
ANI,TensorMol}. ML potentials are much easier to design and 
optimize. With sufficient training data, ML models can achieve similar or even 
better performances compared with density functional theory.
The symmetry function based atomistic neural network (ANN) model, first proposed 
by Parinello and Behler in 2007\cite{Behler,Behler2,Behler3,Behler4,Behler1}, 
is still the most widely used method for \textit{solids} and its effectness and 
accuracy has been proven by several 
works\cite{Behler-solids-lzp-1,Behler-alloy-1,Behler-alloy-2, Behler-alloy-3}.
Until now, various ML implementations\cite{AMP,ANI,TensorMol} have been 
published by researchers. The Amp\cite{AMP} package, developed by Peterson's 
group, is the first open-source framework that implements the symmetry function 
descriptor. Amp can be used to learn interaction potentials for both 
\textit{molecules} and \textit{solids} and it has successfully powered quite a 
few researches since 2016\cite{Amp-works-hammer,amp-work-1}. ANI\cite{ANI} and 
TensorMol\cite{TensorMol} are also symmetry function based packages but they 
mainly focus on \textit{molecules}.

In order to precisely model atomistic interactions, both total energy and atomic 
forces are necessary. For \textit{solids}, the virial stress tensor is also an
essential metric. However, integrating force and stress into the loss function 
is not an easy task. The machine learning approaches favor vectorizable 
operations but the calculations of atomic descriptors are typically complicated 
so that many traditional ML packages (Amp, kCON, etc) choose to pre-compute 
descriptors and only build computation graph from descriptors to total energy. 
This approach can significantly reduce technical difficulty and works very well 
if only total energy is considered in the loss function. But to fit atomic 
forces, numerous codes must be written manually, such as the derivation of 
descriptors with respect to positions. Although the derivation is theoretically
clear, the implementation is challenging and the efficiency can not be 
guaranteed.

Recently, machine learning platforms (TensorFlow\cite{tensorflow}, 
PyTorch\cite{pytorch}, MXNet\cite{MXNet}, etc) have gain significant 
developments. These modern machine learning frameworks can fully take advantage 
of GPUs for acceleration. The automatic differentiation capability brings a new 
route for devloping ML models capable of predicting energy, force and stress as 
we may 'let' ML platforms to handle the calculations of force and stress because 
these metrics can be derived from positions and total energy directly. To 
achieve this, a direct computation graph from positions to total energy must be
built. Thus, the problem becomes: how to design such route?

In this work, we propose a new algorithm, named the virtual atom approach, to 
implement the symmetry function descriptor and the ANN model based on 
TensorFlow. This algorithm is very suitable for parallel execution. With this 
approach, stochastic gradient descent (SGD) based mini-batch training can be 
adopted. We also derive a general and vectorizable expression for computing 
virial stress for arbitrary ANN model. All these algorithms have been 
implemented in our Python program: TensorAlloy.

This paper is organized as follows. Section \ref{section:theory} introduces the 
theory of the symmetry function descriptor and the ANN framework. Section 
\ref{section:method} describes the virtual atom approach and a modified ANN 
model used by TensorAlloy. Section \ref{section:discussions} displays 
performances of TensorAlloy on two public datasets. The appendix 
\ref{section:appendix} gives derivation and implementation details of key 
algorithms. Implementations of the core algorithms and some trained models can 
be obtained from GitHub (https://github.com/Bismarrck/vap) freely.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 2. Theory
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Theory}
\label{section:theory}

\newcommand{\rijn}{r_{ij\mathbf{n}}}

In the atomistic neural network (\textbf{ANN}) framework, the total energy, 
$E^{total}$, of a structure with $N$ atoms, is the sum of all atomic energies:
\begin{equation}
\label{eq:general_e_total}
E^{total} = \sum_{i}^{N}{E_i}
\end{equation}
where $E_{i}$, the energy of atom \textit{i}, is the output of the neural 
network $\mathbf{NN}_{el}$ for element $el$:
\begin{equation}
E_i = \mathbf{NN}_{el}\left( \mathbf{G}_i \right)
\end{equation} 
Here $\mathbf{G}_i$ is a vector representing atomic descriptors of atom $i$. 
$\mathbf{G}_i$ typically only depends on local environments, i.e.  
neighbors of atom $i$:
\begin{equation}
\mathbf{G}_i = F_G(\left\{ \rijn, |\rijn| < r_c\right\} )
\end{equation}
where $F_G$ can be arbitrary function, $r_c$ is the cutoff radius and $\rijn$ is 
the interatomic distance considering the periodic boundary conditions: 
\begin{equation}
    \label{eq:rijn}
    \rijn = \Vert \mathbf{r}_i^{(0)} - \mathbf{r}_j^{(0)} + 
    \mathbf{n}^T \mathbf{h} \Vert
\end{equation}
where $\mathbf{r}_i^{(0)}$ and $\mathbf{r}_j^{(0)}$ denote positions of $i$ and 
$j$ in the primitive cell, $\mathbf{n}$ is a column vector specifying the cell 
shifts along $(X,Y,Z)$ directions 
\begin{equation}
    \label{eq:ij_shift}
    \mathbf{n} = \begin{pmatrix}
        n_x \\
        n_y \\
        n_z
    \end{pmatrix}
\end{equation}
and $\mathbf{h}$ is the \textbf{row-major} $3 \times 3$ lattice tensor:
\begin{equation}
    \label{eq:lattice}
    \mathbf{h} = \begin{pmatrix}
        h_{xx} & h_{xy} & h_{xz} \\
        h_{yx} & h_{yy} & h_{yz} \\
        h_{zx} & h_{zy} & h_{zz} \\
    \end{pmatrix}
\end{equation}

\newcommand{\rijna}{r_{ij\mathbf{n_1}}}
\newcommand{\rikna}{r_{ik\mathbf{n_2}}}
\newcommand{\rjkna}{r_{jk\mathbf{n_3}}}

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor, especially for alloys. The symmetry function descriptor 
consists of two sets of invariant (tranlational, rotational and permutational) 
functions: the radial symmetry function $\mathbf{G}^{(2)}$ and the angular 
symmetry function $\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, \omega) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, \omega)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2(\eta, i, j, \omega) = 
& \exp\left\{-\frac{\eta(\rijn - \omega)^2}{r_c^2} \right\} \cdot f_c(\rijn) \\
\label{eq:g4}
g_4(
    \beta, \gamma, \zeta, 
    i, j, k, 
    \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3) = & 
\left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
\nonumber \\
& \times f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $\omega$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. Equation \eqref{eq:fc_origin} can be further 
transformed to a vectorized expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ should go over all 
neighbors within $r_c$.

Analytical atomic forces can be obtained directly by computing the first-order 
derivative of $E^{total}$ with respect to $\mathbf{r}_{a}^{(0)}$ where $a$ is an 
index:
\begin{equation}
\label{eq:dEdr}
f_a = -\frac{\partial E^{total}}{\partial \mathbf{r}_a^{(0)}}
\end{equation}
According to the chain rule, we can expand Equation \ref{eq:dEdr}:
\begin{align}
\frac{\partial E^{total}}{\partial \mathbf{r}_a^{(0)}} = & \sum_{i}^{N}{
    \frac{
        \partial\mathbf{NN}_{el}(\mathbf{G}_i)}{
        \partial \mathbf{r}_a^{(0)}}} \nonumber \\
= & \sum_{i}^{N}{\sum_{l}^{N_G}{
    \frac{\partial \mathbf{NN}_{el}(\mathbf{G}_i)}{\partial G_{il}}
    \cdot
    \frac{\partial G_{il}}{\partial \mathbf{r}_{a}^{(0)}}
}}
\end{align}
and 
\begin{align}
\frac{\partial G_{il}^{(2)}}{\partial \mathbf{r}_{a}^{(0)}} = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{
        \frac{\partial g_2}{\partial \rijn}
        \cdot
        \frac{\partial \rijn}{\partial r_k^{0}}
    }
} \\
\frac{\partial G_{il}^{(4)}}{\partial \mathbf{r}_{a}^{(0)}} = &
2^{1 - \zeta}\sum_{j, k \neq j, k \neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        \left(
            \frac{\partial g_4}{\partial \rijna}
            \frac{\partial \rijna}{\partial \mathbf{r}_{a}^{(0)}} + 
            \frac{\partial g_4}{\partial \rikna}
            \frac{\partial \rikna}{\partial \mathbf{r}_{a}^{(0)}} + 
            \frac{\partial g_4}{\partial \rjkna}
            \frac{\partial \rjkna}{\partial \mathbf{r}_{a}^{(0)}}
        \right)
    }}}
}
\end{align}

The derivations are theoretically straightforward but their implementations are 
challenging \cite{AMP}. Later we will provide a new approach to address this 
problem.

The $3 \times 3$ virial stress tensor $\epsilon$ is an important metric to 
describe solids under deformation or external pressure. For arbitrary many-body 
interaction potentials, $\epsilon$ can be calculated with the following equation
\cite{lammps_stress}:
\begin{equation}
\label{eq:stress_lammps}
V \cdot \epsilon = \left(-\sum_{i=1}^{N}{\mathbf{r}_i^{(0)} \otimes f_i} - 
\sum_{\mathbf{n}}{\mathbf{h}^T\mathbf{n}} \otimes 
\sum_{i=1}^{N}{F^{\prime}_{i\mathbf{n}}}
\right)^T
\end{equation}
where $V$ is the volume, $\otimes$ indicates tensor product, $f_i$ is the total 
force acting on atom $i$ and $F^{\prime}_{i\mathbf{n}}$ is the partial force. 
Until now, most machine learning force-field models focus on fitting total 
energy and atomic forces, very few of them\cite{DeePMD,DeePMD_rl,DeePMD_kit} had 
attempted to include stress in their loss functions. 
The current expression of Equation \ref{eq:stress_lammps} is not very compatible 
with NN or other ML potentials since the partial forces are difficult to compute 
with vectorized operations. To resolve this issue, we derived an equivalent form 
of Equation \ref{eq:stress_lammps}:
\begin{equation}
\label{eq:stress}
V \cdot \epsilon = -F^{T} R + 
\left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h}
\end{equation}
where $R$ is the $N \times 3$ positions matrix and $F$ is the 
corresponding total forces matrix. The new equation can be easily implemented 
with simple matrix operations. The detailed derivation is given in the appendix.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3. Method
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Method}
\label{section:method}

\newcommand{\nmax}{
    N_{\mathrm{Ni}}^{\mathrm{max}}+N_{\mathrm{Mo}}^{\mathrm{max}}
}
\newcommand{\nijmax}{N_{ij}^{\mathrm{max}}}
\newcommand{\nijkmax}{N_{ijk}^{\mathrm{max}}}
\newcommand{\nelmax}{N^{\mathrm{max}}_{\mathrm{el}}}
\newcommand{\nel}{N_{\mathrm{element}}}
\newcommand{\nvap}{N^{\mathrm{vap}}}
\newcommand{\nnimax}{N^{\mathrm{max}}_{\mathrm{Ni}}}
\newcommand{\nmomax}{N^{\mathrm{max}}_{\mathrm{Mo}}}

In this section we will describe the program design and the core algorithm: 
the virtual atom approach (VAP). The most noticable feature of the TensorAlloy 
program is that the calculation of symmetry function descriptors is finally 
implemented in the compuation graph with the help of VAP. Thus, we can utilize 
the \textbf{AutoGrad} routine of TensorFlow to derive atomic forces and virial 
stress automatically and also efficiently.

The Ni-Mo binary alloy system will be used as the example to show this 
algorithm. For simplicity, only radial symmetry function related algorithms are 
visualized in the figures. The associated codes (Python3.7) are given in the 
appendix. A demo script can be downloaded from GitHub freely 
(https://github.com/Bismarrck/vap).

Fig \ref{fig:program_design} shows the general design of TensorAlloy.
TensorAlloy has two phases: 1) is the training phase and 2) represents the 
prediction phase. These two phases have slightly different focusing points:

In the training phase, efficiency is the major concern. To accelerate model 
training, batching (or mini-batch training) is one of the most commonly used 
techniques. Batching requires aligned inputs. However, hundreds of different 
stoichiometries may exist in one dataset. To batch such structures, an universal 
approach of expressing different structures in a single reference system (the 
global symbol list or GSL) is necessary.

In the prediction phase, things may be a bit different. The stoichiometry of 
the input structure may not be included in the training dataset and it can vary 
significantly. Taking the example of Ni-Mo, the trained model must be capable of 
predicting any $\mathrm{Ni}_{x}\mathrm{Mo}_{y}$ solid while $x$ and $y$ can be 
either as large as 50000 or just zero. Hence, VAP must be flexible enough to 
handle these situations.

In the following subsection, we will discuss the details of VAP.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.1
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.58]{figures/Fig1.pdf}
\caption{\label{fig:program_design} The general design of TensorAlloy. 
\textbf{(1)} demonstrates the training phase while \textbf{(2)} represents the 
prediction phase.
}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Preparation}
\label{section:preparation}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.2
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{figures/Fig2.pdf}
\caption{\label{fig:Fig2} The initialization step. 
\textbf{(1)} describes the detections of $N_{\mathrm{Ni}}^{\mathrm{max}}$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and the global symbol list (GSL). The red 'X' 
represents the inserted virtual atom.
\textbf{(2)} shows the detections of $\nijmax$ and $\nijkmax$ given $r_{cut}$.
}
\end{figure}

Fig \ref{fig:Fig2} shows the initialization of VAP. The cutoff radius $r_{cut}$ 
must be fixed. The initialization involves two steps. The first step is to build 
the global symbol list (GSL), which acts as the universal reference system. A 
virtual atom is inserted at the very first position of this list. The benefits 
will be discussed later. To build GSL, all $\nelmax$ must be obtained. 
$\nelmax$ indicates the maximum appearances of an element in any structure and 
$\nel$ represents the total number of unique chemical elements:
\begin{equation}
\label{eq:n_vap}
\nvap = 1 + \sum_{\mathrm{el}}{\nelmax}
\end{equation}
For the Ni-Mo dataset, $\mathrm{el} \in [\mathrm{Ni}, \mathrm{Mo}]$ and 
$\nel = 2$. 
Table \ref{table:GSL} gives some examples of the GSL mapping. 'X' denotes the 
inserted virtual atom and Ni$_{108(54)}$ means the first 54 Ni atoms are mapped 
from the local stoichiometry. In the training phase, all $\nelmax$ are constant.
In the prediction phase, $\nelmax$ shall be determined dynamically. One should 
note that the minimum value of $\nelmax$ should be 1 due to technical reason (so 
that the computation graph of TensorFlow can be properly executed). The codes to
get $\nelmax$ and GSL are given in Appendix(C). With GSL setup, atomic positions
and forces can be transformed as well. Appendix(D) gives the implementation of 
the virtual-atom mapping mechanism.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table.1
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lccccc}
\hline
 & Stoichiometry & $\nnimax$ & $\nmomax$ & $N^{\mathrm{vap}}$ 
 & Stoichiometry (GSL) \\
\hline
 & Ni$_{54}$ &  &  &  & XNi$_{108(54)}$Mo$_{176(0)}$ \\
Training & Ni$_{54}$Mo$_{12}$ & 108 & 176 & 285 
 & XNi$_{108(54)}$Mo$_{176(12)}$ \\
 & Mo$_{12}$ &  &  &  & XNi$_{108(0)}$Mo$_{176(12)}$ \\
\hline
 & Ni$_{54}$ & 54 & 1 & 56 & XNi$_{54(54)}$Mo$_{1(0)}$ \\
Prediction & Ni$_{54}$Mo$_{12}$ & 54 & 12 & 68 & XNi$_{54(54)}$Mo$_{12(12)}$ \\
 & Mo$_{12}$ & 1 & 12 & 14 & XNi$_{1(0)}$Mo$_{12(12)}$ \\
\hline
\end{tabular}
\caption{\label{table:GSL}
The mappings of local stoichiometries to their GSL forms, taking examples of the 
SNAP/Ni-Mo dataset.
}
\end{table}

Then we can determine another two key constants: $\nijmax$ and $\nijkmax$. 
Here $N_{ij}^{\mathrm{max}}$ represents the maximum number of neighbor pairs in
arbitrary structure and the subscript 'ijk' denotes triples. 
$N_{ij}^{\mathrm{max}}$ and $N_{ijk}^{\mathrm{max}}$ only depend on 
$r_{cut}$ and the associated codes are given in Appendix(B). 
Table \ref{table:nij_nijk_max} summarizes $\nijmax$ and $\nijkmax$ at different
$r_{cut}$ of the two tested datasets: QM7 and SNAP/Ni-Mo.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table.2
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{lccc}
\hline
 Dataset & $r_{cut}$(\AA) & $\nijmax$ & $\nijkmax$ \\
 QM7 & 6.5 & 506 & 5313 \\
\hline
 SNAP/Ni-Mo &  4.6 & 7200 & 176400 \\
 & 6.0 & 12384 & 526320 \\
 & 6.5 & 16284 & 965338 \\
\hline
\hline
\end{tabular}
\caption{\label{table:nij_nijk_max}
$\nijmax$ and $\nijkmax$ of SNAP/Ni-Mo and QM7 at different $r_{cut}$.
}
\end{table}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.B
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Interatomic distances}
\label{section:interatomic_distance}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.3
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{figures/Fig3.pdf}
\caption{\label{fig:Fig3} The pseudo computation graph with execution orders for 
calculating interatomic distances $\mathbf{r}$ from atomic positions 
$\mathbf{R}$ and lattice tensor $\mathbf{h}$.
}
\end{figure}

Both the radial (Equation \ref{eq:g2}) and the angular (Equation \ref{eq:g4}) 
functions only interact with interatomic distances. In order to compute symmetry
function descriptors, local atomic positions shall be transformed to GSL 
interatomic distances first. Fig \ref{fig:Fig3} demonstrates the pseudo 
computation graph for calculating interatomic distances 
$\mathbf{r}^{\mathrm{GSL}}$ for radial descriptors:

\begin{itemize}

    \item[1.] 
    Find all (i,j) neighbor pairs using the routine 
    \href{https://wiki.fysik.dtu.dk/ase/ase/neighborlist.html}
    {\textit{neighbor\_list}} implemented in ASE\cite{ase}. $\mathbf{R}$ 
    represents local atomic positions, $\mathbf{h}$ is the lattice tensor. 
    $\mathbf{i}$ and $\mathbf{j}$ are vectors of length $N_{\mathrm{ij}}$ and
    $\mathbf{n}$ is the $N_{\mathrm{ij}} \times 3$ periodic boundary shift 
    matrix. In the training phase $N_{\mathrm{ij}} \le \nijmax$; in the 
    prediction phase, $\nijmax$ is exactly equal to $N_{\mathrm{ij}}$.

    \item[2.] 
    Map $\mathbf{R}$, $\mathbf{i}$, $\mathbf{j}$ and $\mathbf{n}$ to
    their corresponding GSL forms with padding zeros. One should note that these 
    GSL arrays can be cached for training.

    \item[3.] 
    Broadcast $\mathbf{R}^{\mathrm{GSL}}$ to $\mathbf{R}_{i}^{\mathrm{GSL}}$ and
    $\mathbf{R}_{j}^{\mathrm{GSL}}$ with $\mathbf{i}^{\mathrm{GSL}}$ and 
    $\mathbf{j}^{\mathrm{GSL}}$ respectively.

    \item[4.]
    Compute the displacements in the original cell:
    $\mathbf{D}_{ij} = \mathbf{j}^{\mathrm{GSL}} - \mathbf{i}^{\mathrm{GSL}}$

    \item[5.]
    Compute the total periodic displacements:
    $\mathbf{D}_{pbc} = \mathbf{n}^{\mathrm{GSL}} \mathbf{h}$

    \item[6.]
    Compute the overall GSL displacements:
    $\mathbf{D}^{\mathrm{GSL}} = \mathbf{D}_{pbc} + \mathbf{D}_{ij}$

    \item[7.] 
    Compute the interatomic distances $\mathbf{r}^{\mathrm{GSL}}$.

\end{itemize}

Steps 1-2 are implemented in the function \textit{get\_g2\_map} in Appendix(E).
Steps 3-7 are implemented in the method \textit{get\_rij} of class 
\textit{SymmetryFunction} in Appendix(G).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Radial descriptors}
\label{section:radial_descriptors}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.4
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

Now we can start the last step: computing radial symmetry descriptors 
$\mathbf{G}^{(2)}$ from $\mathbf{r}^{\mathrm{GSL}}$. For simplicity, we will use 
the prediction phase to demonstrate this final section.

Suppose we have $N_{\tau}$ different $(\eta, \omega)$ pairs for Equation 
\ref{eq:G2} and $(\eta, \omega)_{\tau} (0 \le \tau < N_{\tau})$ represents the 
$(\tau+1)$-th pair. In the prediction phase, $\mathbf{r}^{\mathrm{GSL}}$ should 
be a vector of length $\nijmax$ and the desired output $\mathbf{G}^{(2)}$ will 
be a matrix of shape $[\nvap, (\nel)^2 \cdot N_{\tau}]$, as shown in 
Fig \ref{fig:Fig5}\textbf{d}:
\begin{itemize}
    \item[1.] 
    The total number of rows is $\nvap$ (Equation \ref{eq:n_vap}) and each row
    represents the complete radial symmetry function descriptors of an atom. The 
    first row always corresponds to the inserted virtual atom.
    \item[2.]
    The total number of columns is $\nel \cdot \nel \cdot N_{\tau}$. Here 
    $\nel \cdot \nel$ indicates the number of element-element interaction types 
    and each interaction is described by $N_{\tau}$ feature values.
\end{itemize}

\begin{figure}[h!]
\centering
\includegraphics[scale=1.0]{figures/Fig4.pdf}
\caption{\label{fig:Fig4} The visualization of the mapping $f^S_{\nu}$.
In this example $S = [2,4]$ and $p$ can be any $(i,j)$ if 
$0 \le i < 2$ and $0 \le j < 4$.
}
\end{figure}

Assume $\nu$ is a vector of length $N_{\nu}$, we can define an integer matrix 
$f^{S}_{\nu}$ for mapping $\nu$ to a new array $g$ of shape $S$ satisfying the 
following conditions:
\begin{align}
\label{eq:scatter_nd}
% g(p) & = \sum_{\tau}^{}{\nu(\tau)}
g(p) & = \sum_{\tau=0}^{\tau < N_{\nu}}{c_\tau(p) \cdot \nu(\tau)} \\
c_{\tau}(p) & = \begin{cases}
    1 & {f^S}_{\nu}(\tau) = p \\
    0 & \mathrm{else}
\end{cases}
\end{align}
where $p$ is a coordinate vector. The matrix $f^{S}_{\nu}$ shall have $N_{\nu}$ 
rows and $N_S$ columns where $N_S$ represents the dimension of $g$. 
Fig \ref{fig:Fig4} visualizes this $f^S_{\nu}$ mapping scheme. This $f^S_{\nu}$
mapping plays a key role in transforming $\mathbf{r}^{\mathrm{GSL}}$ to 
$\mathbf{G}^{(2)}$.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.5
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.95]{figures/Fig5.pdf}
\caption{\label{fig:Fig5} The definition of $f^{(2,4)}_{\nu}$: map the 
vector $\nu$ to array $g$ of shape [2, 4], or mathematically
$g(i,j) = \sum_{k}^{f^{(2,4)}_{\nu}(k) = (i,j)}{\nu(k)}$.
}
\end{figure}

Fig \ref{fig:Fig5} 
demonstrates the calculation of $\mathbf{G}^{(2)}((\eta, \omega)_{\tau})$ from 
$\mathbf{r}^{\mathrm{GSL}}$. $\mathbf{G}^{(2)}((\eta, \omega)_{\tau})$ is the 
partial descriptors contributed by the $\tau$-th $(\eta, \omega)$ combination.
The overall $\mathbf{G}^{(2)}$ is just the sum of all partial contributions:
\begin{equation}
\mathbf{G}^{(2)} = \sum_{\tau=0}^{\tau < N_{\tau}}{
    \mathbf{G}^{(2)}((\eta, \omega)_{\tau})
}
\end{equation}
This calculation can be easily implemented with a \textit{For} loop. For each 
$\tau$, there will be three major steps:

\begin{itemize}

    \item[1.]
    The first step is applying Equation \ref{eq:g2} on 
    $\mathbf{r}^{\mathrm{GSL}}$ with the given $\tau$-th $(\eta, \omega)$ pair 
    to obtain raw descriptors $g^{\mathrm{GSL}}_{2}((\eta, \omega)_{\tau})$. 
    $g^{\mathrm{GSL}}_{2}((\eta, \omega)_{\tau})$ is a vector of length 
    $\nijmax$ and its last ($\nijmax - N_{ij}$) components are dummy values.

    \item[2.]
    The second step is generating the mapping $f_{g,\tau}^{S}$. In the Ni-Mo 
    case, $S=[285, 4N_{\tau}]$. $f_{g,\tau}^{S}$ is a matrix of shape 
    $[\nijmax, 2]$ and $f_{g,\tau}^{S}$ is constructed from the base mapping 
    array $f_{g}^{S}$ by adding $\tau$ to its second column. $f_{g}^{S}$ is also
    a $\nijmax \times 2$ matrix. $f_{g}^{S}$ stores fundamental properties of 
    $\mathbf{r}^{\mathrm{GSL}}$: the center atom (e.g. $\mathrm{Ni}_{1}$) and 
    type (e.g. Ni-Mo) of each atom-atom interaction:
    \begin{itemize}

        \item[a.] 
        Its first column is directly copied $i^{\mathrm{GSL}}$, representing the 
        GSL indices of the center atoms.

        \item[b.] 
        The values of its second column are multiples of $N_{\tau}$ indicating 
        the offsets, or horizontal positions in $\mathbf{G}^{(2)}$. These values 
        are obtained by multiplying $N_{\tau}$ with the type indices of 
        $\mathbf{r}^{\mathrm{GSL}}$. For example, the Ni-Mo dataset has two 
        unique elements: Ni, Mo. Hence, there will be four types of interactions:
        Mo-Mo (0), Mo-Ni (1), Ni-Ni (2) and Ni-Mo (3). Then for each value 
        (atom-atom interaction) of $\mathbf{r}^{\mathrm{GSL}}$, its type index 
        can be determined with $\mathbf{i}^{\mathrm{GSL}}$ and 
        $\mathbf{j}^{\mathrm{GSL}}$ based on the pre-built GSL.

    \end{itemize}

    \item[3.]
    Compute $\mathbf{G}^{(2)}((\eta, \omega)_{\tau})$ by applying 
    $f_{g,\tau}^{S}$ on $g^{\mathrm{GSL}}_{2}((\eta, \omega)_{\tau})$. The 
    padded zeros in $\mathbf{r}^{\mathrm{GSL}}$ are finally mapped to the 
    inserted virtual atom. Hence, the zero-padding of 
    $\mathbf{r}^{\mathrm{GSL}}$ can be easily separated and elinimated (just 
    remove the first row of $\mathbf{G}^{(2)}$). That's why TensorAlloy can use
    arbitrary number of structures without any stoichiometry restriction to do 
    batch training.

\end{itemize}

The mapping scheme $f_{\nu}^{S}$ can be directly achieved using routine 
\textit{tf.scatter\_nd} implemented by TensorFlow. The calculation of the base 
mapping array $f_{g}^{S}$ is implemented in Appendix(E). $f_{g}^{S}$ is also 
cacheable. The calculation of $\mathbf{G}^{(2)}((\eta, \omega)_{\tau})$ is 
implemented in method \textit{get\_g2\_op\_for\_tau} of class 
\textit{SymmetryFunction} in Appendix(G).

The workflow for the batch training phase is almost identical to the prediction 
phase introduced above. There are still some noticable diiferences. Assume 
$N_{b}$ represents the batch size:

\begin{itemize}

    \item[1.] 
    Most of the variables ($\mathbf{r}^{\mathrm{GSL}}$, 
    $\mathbf{i}^{\mathrm{GSL}}$, $\mathbf{R}^{\mathrm{GSL}}$, etc.) have one 
    more dimension. For example, $\mathbf{R}^{\mathrm{GSL}}$ should be an array
    of shape $[N_b, \nijmax, 3]$.

    \item[2.] 
    The base mapping $f_{g}^{S}$ becomes a matrix of shape $[\nijmax, 3]$. Its 
    last two columns remain unchanged. However, the newly inserted first column 
    \textemdash representing the batch indices \textemdash is 
    \textbf{not cacheable} because batching is a random action. The first column
    should be generated dynamically.

    \item[3.]
    $\tau$ should be added to the third column of $f_{g}^{S}$.

    \item[4.]
    In the training phase, $\nvap$ is a constant determined before training. In 
    the contrast, $\nvap$ becomes a tensor that should be manually specified in 
    the prediction phase.

    \item[5.]
    In the training phase, the reference atomic forces should be transformed to 
    its corresponding GSL array because the NN-derived force array 
    $\mathbf{F}^{\mathrm{NN}}$ is in GSL reference:
    \begin{equation}
        \mathbf{F}^{\mathrm{NN}} = 
        -\frac{\partial{E}}{\partial{\mathbf{R}^{\mathrm{GSL}}}}
    \end{equation}
    Hence, in the prediction phase, a reverse VAP mapping is required to convert
    $\mathbf{F}^{\mathrm{NN}}$ to its local form.

\end{itemize}

The training phase is implemented in \textit{BatchSymmetryFunction} which 
inherits most of methods from its parent class \textit{SymmetryFunction}. 
The dynamic batching is implemented in method 
\textit{get\_v2g\_map\_batch\_indexing\_matrix}. The method 
\textit{get\_g\_shape} will return a fixed array instead. The function
\textit{map\_gsl\_array\_to\_local} of \textit{VirtualAtomMap} is used to do the 
reverse mapping.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.D
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Angular descriptors}

The same algorithm can be applied to compute angular symmetry function 
descriptors. The triple list shall be constructed from the neighbor list 
(Fig \ref{fig:Fig3}) and $\nijmax$ shall be replaced by $\nijkmax$. Typically
$\nijkmax \gg \nijmax$.

Appendix(F) demonstrates how to compute the base mapping array. The calculation 
of $\mathbf{G}^{(4)}(\beta, \gamma, \zeta)_{\tau}$ is implemented in method 
\textit{get\_g4\_op\_for\_tau} of class \textit{SymmetryFunction} 
in Appendix(G).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.E
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Min-max normalization}

Raw values of $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ may range from zero to 
several hundreds. So a min-max normalization routine is recommended. For each
column $\mathbf{G}(:, \tau)$ where $0 \le \tau < N_{\tau}$, we just normalize 
its values to range [0, 1]:
\begin{equation}
    \mathbf{G}(:, \tau) = \frac{
        \mathbf{G}(:, \tau) - \min(\mathbf{G}(:, \tau))}{
        \max(\mathbf{G}(:, \tau)) - \min(\mathbf{G}(:, \tau))}
\end{equation}
where $\min(\mathbf{G}(:, \tau))$ and $\max(\mathbf{G}(:, \tau))$ should be 
determined in the training phase and remain constant in the prediction phase.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.F
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{The residual model}

To fit the total energy, we slightly modified the total energy expression of 
Equation \ref{eq:general_e_total}:
\begin{align}
\label{eq:e_total_res}
E^{total} & = E^{residual} + E^{static} \nonumber \\
& = \sum_{\mathrm{el}}{\mathbf{NN}_{el}\left(\mathbf{G}_{\mathrm{el}}\right)}
+ \sum_{\mathrm{el}}{n_{el}E_{el}}
\end{align}
where $\{E_{el}\}$ is a set of \textit{trainable} scalar variables representing 
the \textit{static} energy of each type of element\cite{kCON}. The NN in 
Equation \ref{eq:e_total_res} just describes the atomistic interactions 
(the \textit{residual} energy). 
The introduction of $E^{static}$ can significantly limit value range of NN 
outputs, leading to faster convergence speed and higher training stability 
(Fig \ref{fig:benchmark_qm7}c). The initial $\{E_{el}\}$ are calculated by 
solving the linear system $Ax=b$ where $A$ is a $N_{data} \times \nel$ matrix 
and $b$ is a vector of length $N_{data}$. $N_{data}$ is the number of training 
examples. $A(i,j)$ is the number of $j$-th element in structure $i$ and $b(i)$ 
is the corresponding total energy.

To train the model, we use the following root mean-squared error (RMSE) total 
loss function:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3 N_{b} \nvap}\sum_{i}^{N_b}{\sum_{j}^{\nvap}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{s}}\sqrt{\frac{1}{6N_b}\sum_{i}^{N_b}{
    \sum_{j}^{6}{
        \left(
            \epsilon^{\mathrm{voigt}}_{j} - \epsilon^{\mathrm{voigt,dft}}_{j}
        \right)^2
    }
}}
\end{align}
where $N_b$ is the mini-batch size and $\chi_{\mathrm{f}}$ and 
$\chi_{\mathrm{s}}$ are weights of force and stress losses. 
Stress tensors are converted to Voigt vectors. All energies are in \textbf{eV}, 
all forces are in \textbf{eV/\AA} and all stress components are in 
\textbf{eV/\AA}$\mathbf{^3}$. In most cases, we set $N_b$ to 50 or 100 and we
use the ADAM\cite{adam} optimizer with exponentially-decayed learning rate to 
minimize the loss function. The default activation function is softplus:
\begin{equation}
\sigma(x) = \log(1 + e^x)
\end{equation}
The kernel weights of all NNs are initialized with the 
Xvaier\cite{pmlr-v9-glorot10a} method and kernel biases are initialized with 
zeros.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4. Discussions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Discussions}
\label{section:discussions}

In this section we demonstrate two experiments of TensorAlloy. All GPU benchmark
results are obtained on the same workstation with two Intel Xeon E5-2687v4 CPUs
(18 cores @ 2.3 GHz per CPU) and one NVIDIA GTX 1080Ti GPU.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{QM7}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.6
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
\centering
\includegraphics[scale=0.54]{figures/qm7_speed.pdf}
\caption{\label{fig:benchmark_qm7} 
\textbf{a)} The curves of training speed (structures per second) vs batch sizes. 
\textbf{b)} The curves of test MAE (meV/atom) vs training time (GPU hour) using 
different settings.
\textbf{c)} The curves of test MAE (meV/atom) vs NN models. 
}
\end{figure}

QM7\cite{QM7_1,QM7_2} is a publicly available benchmark dataset calculated at 
the DFT level. The QM7 dataset contains 7165 stable organic molecules 
(C, H, N, O, S) with 176 unique stoichiometries. The total energies range from 
-95 eV to -17 eV and the structure sizes range from 5 to 23. 1000 structures 
were randomly selected as test set before training.

Figure \ref{fig:benchmark_qm7} demonstrates the performances of 
TensorAlloy on QM7 dataset. In this figure, 'Radial' means only radial 
symmetry functions (Equation \ref{eq:G2}) are used and 'Angular' indicates both 
radial and angular symmetry functions are used. The cutoff radius is 6.5 \AA.
The corresponding $\nijmax$ and $\nijkmax$ are 506 and 5313, respectively.
Values of $\eta$ for radial symmetry functions are 0.1, 0.5, 1, 2, 4, 8, 12, 16, 
20, 40 and $\beta$ (0.1, 0.5), $\gamma$ (1, -1) and $\zeta$ (1, 4) are used for
angular functions. $\omega$ is fixed to 0. The initial learning rate is 0.01. 
Learning rate decay is disabled for the following tests. Min-max normalization
is enabled. Each atomistic neural network has two hidden layers with 64 and 32 
neurons. The force and stress losses of Equation \ref{eq:loss} are disabled for 
this experiment.

Figure \ref{fig:benchmark_qm7}a compares the training speed (number of processed 
structures per second) with different batch sizes (number of structures per 
step). It shows that CPU perfers fewer descriptors (radial only) and smallber 
batch size while GPU is suitable for larger batch size and more complicated 
descriptors (radial + angular). 
Figure \ref{fig:benchmark_qm7}b shows the curves of mean absolute errors 
(MAEs, meV/atom) on the test set. This figure clearly illustrate another benefit 
of using larger batch size: the trainings can be much faster. For the best case 
(batch size 100, angular symmetry functions included), the test MAE will reach 5 
meV/atom (1.5 kcal/mol per structure) in just one GPU hour.

Figure \ref{fig:benchmark_qm7}c compares out residual model (Equation 
\ref{eq:e_total_res}) with traditional ANN model (Equation 
\ref{eq:general_e_total}). The proposed residual model has a much faster 
convergence rate and better stability. One exaplanation is that the 
\textit{static} part acts as a normalization function. By introducing the 
\textit{static} term, the output values of ANNs are much closer to the ideal 
output range (-1, 1) of neural networks.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4.B
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Ni-Mo}

SNAP/Ni-Mo\cite{SNAP_Mo_2017, SNAP_2018} is a publicly available dataset built 
by Shyue Ping Ong and co-workers. This dataset has 3971 Ni-Mo solids, including 
461 pure Ni structures and 284 pure Mo structures. All DFT calculations were 
done by VASP\cite{VASP} using the PBE\cite{PBE} functional within the projector 
angumented-wave (PAW)\cite{PAW} approach.

In this experiment, we trained three different models:
\begin{enumerate}

    \item Ni: this model uses 400 Ni structures for training and 61 for 
    evaluation. The loss function only includes energy and force contributions.

    \item Mo: this model uses 250 Mo structures for training and 34 for 
    evaluation. The loss function includes all three (energy, force, stress) 
    types of contributions.

    \item Ni-Mo: this model uses 3673 structures for training and 300 for 
    evaluation. The loss function includes all three types of contributions.

\end{enumerate}

For all these models, the cutoff radius is set to 6.5 \AA. Only radial symmetry 
functions are used to compute atomic descriptors. The selected $\eta$ are 
0.1, 0.5, 1, 2, 4, 8, 12, 16, 20, 40 and $\omega$ are 0.0, 3.0. The hidden layer
sizes are 128, 64, 32. The activation function is softplus.
The learning rate starts from 0.01 and it will decay exponentially with rate 
0.95 for every 3000 steps.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table. 3
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}[h]
\centering
\begin{tabular}{cccc|ccc|cc}
\hline
 & \multicolumn{3}{c|}{Energy (meV/atom)} & \multicolumn{3}{c|}{Force (eV/\AA)}
 & \multicolumn{2}{c}{Stress (GPa)} \\
\hline
Model & Mo & Ni & Ni-Mo & Mo & Ni & Ni-Mo & Mo & Ni-Mo \\
\hline
Mo & 4.9 (13.2) & & 16.5 (16.2) & 0.19 (0.25) & & 0.28 (0.29) & 0.24 (0.87) 
   & 0.75 \\
Ni & & 1.2 (1.2) & 4.5 (7.9) & & 0.04 (0.05) & 0.07 (0.11) & & 0.83 \\
Ni$_4$Mo & & & 4.1 (4.0) & & & 0.09 (0.14) & & 0.98 \\ 
Ni$_3$Mo & & & 4.5 (5.2) & & & 0.11 (0.16) & & 1.19 \\
Ni$_{\mathrm{Mo}}$ & & & 12.9 (22.7) & & & 0.09 (0.13) & & 0.34 \\
Mo$_{\mathrm{Ni}}$ & & & 12.9 (33.9) & & & 0.12 (0.55) & & 0.45 \\
Overall & & & 10.8 (22.5) & & & 0.11 (0.23) & & 0.59 \\
\hline
\end{tabular}

\caption{\label{table:MAE}
Comparion of the MAEs in predicted energies (mev/atom), forces (eV/\AA) and 
stress (GPa) relative to DFT for TensorAlloy-trained models and their
corresponding SNAP benchmarks (bracket)}
\end{table}

Table \ref{table:MAE} summarizes the energy, force and stress prediction 
performances of TensorAlloy-trained models compared with their corresponding 
SNAP benchmarks. The original SNAP formalism contains both radial and angular 
interactions. However, in most cases, the radial-only TensorAlloy-trained 
models can significantly outperform corresponding SNAP models. The trainings of 
these models are also very fast. All these models are trained within 2 GPU hours
on our workstation.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.7
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.64]{figures/Prediction-speed.pdf}
\caption{\label{fig:prediction_speed} Benchmarks of the Ni-Mo model in 
the prediction phase. The test structures are (MoNi)$_{x}$ where 
$x$ = $10^3, 15^3, 20^3, 25^3, 30^3, 35^3, 40^3$. The left figure \textbf{a)} 
demonstrates the overall time and the right figure \textbf{b)} compares the 
'Graph' time with/out GPU. Blue bars indicate the elapsed time (in seconds) for 
finding neighbors, orange bars denote the time for initializing VAP-GSL arrays 
and green bars and the numbers above represents the graph execution time (used 
for obtaining energy, forces and stress).
}
\end{figure}

The speed of the prediction phase is also a crucial metric for measuring 
interaction potentials. Fig \ref{fig:prediction_speed} summarizes the time 
distribution of the Ni-Mo model in the prediction phase on the GPU workstation. 
Here 'prediction' is defined as calculating energy, forces and stress tensor of 
an arbitrary \textit{ase.Atoms} object. In these tests, the primitive structure
$\mathrm{MoNi}$ is obtained by replacing one of the two bcc Mo atoms with a Ni 
atom. The test subjects $(\mathrm{MoNi}))_{x}$ are just supercells of the base 
structure. We choose $x \in [10^3, 15^3, 20^3, 25^3, 30^3, 35^3, 40^3]$ as
a typical molecular dynamics simulation for studying alloy properties needs at 
least thousands of atoms. The elaspsed time is splitted to three parts: 
'Neighbor' (blue) indicates the execution time of the \textit{neighbor\_list} 
routine of ASE for finding local neighbors (\textbf{(1)} of Fig \ref{fig:Fig3}),
'VAP' (orange) denotes the time to construct VAP-GSL arrays 
(\textbf{(2)} of Fig \ref{fig:Fig3}) and 'Graph' (green bars and the numbers 
above) represents the time of getting energy, forces and stress tensor by 
executing the computation graph. The smallest subject (2000 atoms) needs 1.6 
seconds and the largest system (128000 atoms) requires ~71.6 seconds on the 
workstation. Most of the time are spent on the Python-based single-thread 
'Neighbor' and 'VAP' routines \textemdash which may be greatly optimized by 
utilizing efficient C/C++/Fortran codes (e.g. Lammps). For the 'Graph' part, the 
execution time can be as small as 1 second (with GPU) or 4 seconds (without GPU) 
for the 128000-atoms structure. According to these tests, it's fairly practical 
to use TensorAlloy-trained symmetry function interaction potential on studying 
large alloy systems.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.8
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.8]{figures/Ni_EoS.pdf}
\caption{\label{fig:energy_volume_Ni} The energy vs volume curves of fcc Ni for
the DFT, Ni and Ni-Mo models.}
\end{figure}

To further validate our models, we also calculate the energy-volume curves of a 
conventional fcc Ni cell using SFNN/Ni and SFNN/Ni-Mo models. The results are 
plotted in Fig \ref{fig:energy_volume_Ni}. Both curves overlap the DFT curve 
very well in the range of -17\% to 21\% from the equilibrium volume. The surface
energy tests, as shown in Fig \ref{fig:surface_energy_Ni}, also proves the 
accuracy of TensorAlloy-optimized models. PyMatGen\cite{pymatgen,pymatgen-1} is 
used to generate these surface slabs.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Fig.9
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.8]{figures/Ni_surface_energy.pdf}
\caption{\label{fig:surface_energy_Ni} Surface formation energies ($J/m^2$) of 
different Ni surfaces calculated by DFT, Ni and Ni-Mo models}
\end{figure}

The benchmark function and the trained models can be accessed from the VAP 
GitHub repo (https://github.com/Bismarrck/vap) freely.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 5. Conclusions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Conclusions}
\label{section:conclusions}

In this work, we proposed a two-phases (training, prediction) approach 
(the virtual atom approach) to construct symmetry function based atomistic 
neural networks. 
With this method, we can build direct computation graph from atomic positions to 
total energy within TensorFlow, thus making the technical barrier of calculating 
NN-derived atomic forces and virial stress neglectable. We also derived an 
alternative form of the general virial stress equation. This equation can be 
calculated with simple matrix operations, which makes it suitable for machine 
learning based interaction potentials.
With these new features, we developed a Python program, TensorAlloy. This 
program can be used to learn interaction potentials of \textit{molecules} or 
\textit{solids}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Acknowledgments
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section*{Acknowledgments}

This work was supported by the National Key Research and Development Program of 
China under Grant No. 2016YFB0201203, the Science Challenge Project under Grant 
No. TZ2018002 and the National Natural Science Foundation of China under Grant 
No. U1630250.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% References
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\bibliography{manuscript.bib}

\newpage

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Appendix
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section*{Appendix}
\label{section:appendix}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section Appendix. A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Derivation and implementation of the virial stress equation}

According to Equation \ref{eq:lattice} and Equation \ref{eq:ij_shift}, Equation 
\ref{eq:rijn} can be expanded:
\begin{align}
\rijn = & \Vert \mathbf{r}_i^{(0)} - \mathbf{r}_j^{(0)} + 
           \mathbf{n}^T \mathbf{h} \Vert \nonumber \\
      = & \left[
                 \left(r_{j,x}^{(0)} - r_{i,x}^{(0)} + 
                       \sum_{\alpha}{n_{\alpha}h_{\alpha x}} \right)^2 +
                 \left(r_{j,y}^{(0)} - r_{i,y}^{(0)} + 
                       \sum_{\alpha}{n_{\alpha}h_{\alpha y}} \right)^2 + \right.
        \nonumber \\
        & \left. \left(r_{j,z}^{(0)} - r_{i,z}^{(0)} + 
                       \sum_{\alpha}{n_{\alpha}h_{\alpha z}} \right)^2 
          \right]^{\frac{1}{2}}
\end{align}
where $\alpha=x,y,z$. Thus, we can calculate the derivation of $\rijn$ with 
respect to $h_{\alpha\beta}$:
\begin{gather}
\frac{\partial \rijn}{\partial h_{\alpha\beta}} = 
    \frac{1}{\rijn} \cdot \Delta_{ij\mathbf{n}\beta} \cdot n_{\alpha} \\
\Delta_{ij\mathbf{n}\beta} = r_{j,\beta}^{(0)} - r_{i,\beta}^{(0)} + 
    \sum_{\alpha}{n_{\alpha}h_{\alpha \beta}}
\end{gather}

\newcommand{\hab}{h_{\alpha\beta}}
\newcommand{\hga}{h_{\gamma\alpha}}
\newcommand{\hgb}{h_{\gamma\beta}}

Then we can derive $\partial E^{total} / \partial \hab$:
\begin{align}
\label{eq:dEdhab}
\frac{\partial E^{total}}{\partial \hab} = \sum_{i}^{N}{\sum_{l}^{N_G}}{
    \frac{\partial \mathbf{NN}_{el}(\mathbf{G}_i)}{\partial G_{il}}
    \cdot
    \frac{\partial G_{il}}{\partial \hab}
}
\end{align}
If $G_{il}$ is a radial symmetry function:
\begin{align}
\frac{\partial G_{il}^{(2)}}{\partial \hab} 
& = \sum_{j \neq i}{\sum_{\mathbf{n}}}{
    \frac{\partial g_2}{\partial \rijn} 
    \cdot 
    \frac{1}{\rijn} \cdot \Delta_{ij\mathbf{n}\beta} \cdot n_{\alpha}
} \\
\label{eq:stress_g2_right}
\frac{\partial G_{il}^{(2)}}{\partial \hga} \cdot \hgb & = 
\sum_{j \neq i}{\sum_{\mathbf{n}}}{
    \frac{\partial g_2}{\partial \rijn} 
    \cdot 
    \frac{\Delta_{ij\mathbf{n}\alpha}}{\rijn} \cdot n_{\gamma}
} \cdot \hgb
\end{align}
If $G_{il}$ is an angular symmetry function:
\begin{align}
\frac{\partial G_{il}^{(4)}}{\partial \hab} & =
    \sum_{j, k \neq j, k \neq i}{
        \sum_{\mathbf{n_1}}}{\sum_{\mathbf{n_2}}{\sum_{\mathbf{n_3}}{
    \left(
        \frac{\partial g_4}{\partial \rijna}\cdot\frac{\partial \rijna}{\hab} +
        \frac{\partial g_4}{\partial \rikna}\cdot\frac{\partial \rikna}{\hab} +
        \frac{\partial g_4}{\partial \rjkna}\cdot\frac{\partial \rjkna}{\hab}
    \right)
}}} \nonumber \\
& =  \sum_{j, k \neq j, k \neq i}{
        \sum_{\mathbf{n_1}}}{\sum_{\mathbf{n_2}}{\sum_{\mathbf{n_3}}{
    \frac{\partial g_4}{\partial \rijna} \cdot 
    \frac{\Delta_{ij\mathbf{n_1}\beta}}{\rijna} \cdot n_{1,\alpha}
}}} \nonumber \\
& + \sum_{j, k \neq j, k \neq i}{
    \sum_{\mathbf{n_1}}}{\sum_{\mathbf{n_2}}{\sum_{\mathbf{n_3}}{
\frac{\partial g_4}{\partial \rikna} \cdot 
\frac{\Delta_{ik\mathbf{n_2}\beta}}{\rikna} \cdot n_{2,\alpha}
}}} \nonumber \\
& + \sum_{j, k \neq j, k \neq i}{
    \sum_{\mathbf{n_1}}}{\sum_{\mathbf{n_2}}{\sum_{\mathbf{n_3}}{
\frac{\partial g_4}{\partial \rjkna} \cdot 
\frac{\Delta_{jk\mathbf{n_2}\beta}}{\rjkna} \cdot n_{3,\alpha}
}}} \\
\label{eq:stress_g4_right}
\frac{\partial G_{il}^{(4)}}{\partial \hga} \cdot \hgb & = 
\sum_{j, k \neq j, k \neq i}{
    \sum_{\mathbf{n_1}}}{\sum_{\mathbf{n_2}}{\sum_{\mathbf{n_3}}{
\frac{\partial g_4}{\partial \rijna} \cdot 
\frac{\Delta_{ij\mathbf{n_1}\alpha}}{\rijna} \cdot n_{1,\gamma} \cdot \hgb
}}} \nonumber \\
& + \sum_{j, k \neq j, k \neq i}{
    \sum_{\mathbf{n_1}}}{\sum_{\mathbf{n_2}}{\sum_{\mathbf{n_3}}{
\frac{\partial g_4}{\partial \rikna} \cdot 
\frac{\Delta_{ik\mathbf{n_1}\alpha}}{\rikna} \cdot n_{2,\gamma} \cdot \hgb
}}} \nonumber \\
& + \sum_{j, k \neq j, k \neq i}{
    \sum_{\mathbf{n_1}}}{\sum_{\mathbf{n_2}}{\sum_{\mathbf{n_3}}{
\frac{\partial g_4}{\partial \rjkna} \cdot 
\frac{\Delta_{jk\mathbf{n_1}\alpha}}{\rjkna} \cdot n_{3,\gamma} \cdot \hgb
}}}
\end{align}
Combining with Equation \ref{eq:dEdhab}, \ref{eq:stress_g2_right} and 
\ref{eq:stress_g4_right}, we can now start deriving the right part of Equation 
\ref{eq:stress}. To simplify the derivation, we just assume $G_{il}$ represents
a radial function:
\newcommand{\dE}{\partial{E^{total}}}
\begin{align}
\label{eq:dEdhTh_g2_expanded}
\left(
    \left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h}
\right)_{\alpha\beta} & = 
\sum_{\gamma}{\frac{\dE}{\partial \hga} \hgb} \nonumber \\
& = \sum_{\gamma}{\sum_{i}^{N}{\sum_{l}^{N_G}}{
    \frac{\partial \mathbf{NN}_{el}(\mathbf{G}_i)}{\partial G_{il}}
    \cdot
    \sum_{j \neq i}{\sum_{\mathbf{n}}}{
    \frac{\partial g_2}{\partial \rijn} 
    \cdot 
    \frac{\Delta_{ij\mathbf{n}\alpha}}{\rijn} \cdot n_{\gamma}} \hgb
}} \nonumber \\
& = \sum_{i}^{N}{\sum_{l}^{N_G}}{\sum_{j \neq i}{\sum_{\mathbf{n}}}{
    \frac{\partial \mathbf{NN}_{el}(\mathbf{G}_i)}{\partial G_{il}}
    \cdot
    \frac{\partial g_2}{\partial \rijn} 
    \cdot 
    \frac{\Delta_{ij\mathbf{n}\alpha}}{\rijn}}
    \sum_{\gamma}{\cdot n_{\gamma}\hgb}
} \nonumber \\
& = -\sum_{i}^{N}{\sum_{l}^{N_G}}{\sum_{j \neq i}{\sum_{\mathbf{n}}}{
    f^{'}_{ij\mathbf{n}\alpha}
    \sum_{\gamma}{\cdot n_{\gamma}\hgb}
}}
\end{align}
where $f^{'}_{ij\mathbf{n}}$ is the partial force:
\begin{equation}
f^{'}_{ij\mathbf{n}\alpha} = 
-\frac{\partial \mathbf{NN}_{el}(\mathbf{G}_i)}{\partial G_{il}} \cdot 
\frac{\partial g_2}{\partial \rijn} \cdot 
\frac{\Delta_{ij\mathbf{n}\alpha}}{\rijn}
\end{equation}
If $G_{il}$ is an angular symmetry function, we can also get a similar 
expression. Thus,
\begin{equation}
\left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h} =
-\left(\sum_{\mathbf{n}}{\mathbf{h}^T\mathbf{n}} \otimes 
\sum_{i=1}^{N}{F^{\prime}_{i\mathbf{n}}}\right)^T
\end{equation}
The left part of Equation \ref{eq:stress} can be calculated with 
simple matrix multiplacation:
\begin{equation}
\sum_{i=1}^{N}{\mathbf{r}_i^{(0)} \otimes f_i} = R^{T} F
\end{equation}
So finally we can derive the vectorized expression of virial stress:
\begin{equation}
\epsilon = -F^{T} R + \left( \frac{\dE}{\partial \mathbf{h}} \right)^T\mathbf{h}
\end{equation}

Now, the virial stress can be calculated with just a few lines of codes within 
arbitrary Machine Learning framework (TensorFlow, PyTorch, etc). The following 
codes are written in Python-3.7/TensorFlow-1.12, where \textit{energy} and 
\textit{volume} are scalar tensors, \textit{cell} is a $3 \times 3$ tensor, 
\textit{positions} is a $N\times 3$ tensor and \textit{forces} (the total forces 
on these atoms) is also a $N\times 3$ tensor:
\begin{verbatim}
def get_virial_stress_tensor(energy, cell, volume, positions, forces):
    dEdh = tf.gradients(energy, cell, name='dEdh)[0]
    right = tf.matmul(tf.transpose(dEdh, name='dEdhT'), cell)
    left = tf.matmul(tf.transpose(forces), positions)
    stress = tf.add(tf.negative(left), right, name='stress')
    stress = tf.div(stress, volume, name='virial')
    return stress
\end{verbatim}

\newpage

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section Appendix. B
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{
    Calculations of \texorpdfstring{$N_{ij}^{\mathrm{max}}$}{Nijmax} and 
    \texorpdfstring{$N_{ijk}^{\mathrm{max}}$}{Nijkmax}
}

The following function returns $N_{ij}$ and $N_{ijk}$ of an \textmd{ase.Atoms}:
\begin{verbatim}
def get_nij_and_nijk(atoms, rc, angular=False):
    ilist, jlist = neighbor_list('ij', atoms, cutoff=rc)
    nij = len(ilist)
    if angular:
        nl = {}
        for i, atomi in enumerate(ilist):
            if atomi not in nl:
                nl[atomi] = []
            nl[atomi].append(jlist[i])
        nijk = 0
        for atomi, nlist in nl.items():
            n = len(nlist)
            nijk += (n - 1 + 1) * (n - 1) // 2
    else:
        nijk = 0
    return nij, nijk
\end{verbatim}
During the training phase, $\nijmax$ and $\nijkmax$ of a dataset (a collection 
of \textmd{ase.Atoms} objects) can be pre-computed with the following codes:
\begin{verbatim}
nij_max = 0
nijk_max = 0
for atoms in dataset:
    nij, nijk = get_nij_and_nijk(atoms, rc, angular)
    nij_max = max(nij, nij_max)
    nijk_max = max(nijk, nijk_max)
\end{verbatim}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section Appendix. C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{
    Calculations of 
    \texorpdfstring{$N_{\mathrm{element}}^{\mathrm{max}}$}{Nelmax} 
}

The following function returns all $N_{\mathrm{element}}^{\mathrm{max}}$ (e.g., 
$N_{\mathrm{Ni}}^{\mathrm{max}}$) of a dataset:
\begin{verbatim}
def get_max_occurs(dataset):
    from collections import Counter
    max_occurs = Counter()
    for atoms in dataset:
        c = Counter(atoms.get_chemical_symbols())
        for e, n in c.items():
            max_occurs[e] = max(n, max_occurs[e])
    return max_occurs
\end{verbatim}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section Appendix. D
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Implementation of the Virtual-Atom Map}

The following class implements the core of the Virtual-Atom Approach. The 
argument \textit{max\_occurs} just represents the Global Symbol List (GSL). The
positions and forces can be transformed to their GSL forms by using the function
\textit{map\_to\_gsl\_array}.

\begin{verbatim}
class VirtualAtomMap:
    REAL_ATOM_START = 1

    def __init__(self, max_occurs: Counter, symbols: List[str]):
        istart = VirtualAtomMap.REAL_ATOM_START
        self.max_occurs = max_occurs
        self.symbols = symbols
        self.max_vap_n_atoms = sum(max_occurs.values()) + istart
        elements = sorted(max_occurs.keys())
        offsets = np.cumsum([max_occurs[e] for e in elements])[:-1]
        offsets = np.insert(offsets, 0, 0)
        delta = Counter()
        index_map = {}
        mask = np.zeros(self.max_vap_n_atoms, dtype=bool)
        for i, symbol in enumerate(symbols):
            i_ele = elements.index(symbol)
            i_old = i + istart
            i_new = offsets[i_ele] + delta[symbol] + istart
            index_map[i_old] = i_new
            delta[symbol] += 1
            mask[i_new] = True
        reverse_map = {v: k - 1 for k, v in index_map.items()}
        index_map[0] = 0
        reverse_map[0] = -1
        self.atom_masks = mask
        self.index_map = index_map
        self.reverse_map = reverse_map
        self.splits = np.array([1, ] + [max_occurs[e] for e in elements])

    def map_array_to_gsl(self, array: np.ndarray):
        rank = np.ndim(array)
        if rank == 2:
            array = array[np.newaxis, ...]
        elif rank <= 1 or rank > 3:
            raise ValueError("The rank should be 2 or 3")
        if array.shape[1] == len(self.symbols):
            array = np.insert(
                array, 0, np.asarray(0, dtype=array.dtype), axis=1)
        else:
            shape = (array.shape[0], len(self.symbols), array.shape[2])
            raise ValueError(f"The shape should be {shape}")
        indices = []
        for i in range(self.max_vap_n_atoms):
            indices.append(
                self.reverse_map.get(i, -1) + self.REAL_ATOM_START)
        output = array[:, indices]
        if rank == 2:
            output = np.squeeze(output, axis=0)
        return output

    def reverse_map_gsl_to_local(self, array: np.ndarray):
        rank = np.ndim(array)
        if rank == 2:
            array = array[np.newaxis, ...]
        assert array.shape[1] == self.max_vap_n_atoms
        istart = self.REAL_ATOM_START
        indices = []
        for i in range(istart, istart + len(self.symbols)):
            indices.append(self.index_map[i])
        output = array[:, indices]
        if rank == 2:
            output = np.squeeze(output, axis=0)
        return output
\end{verbatim}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section Appendix. E
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Implementation of the \texorpdfstring{$g^2_{\mathrm{map}}$}{g2map}}

This function shows the calculation of $g^2_{\mathrm{map}}$. The returned dict 
is organized as follows: 
\begin{itemize}
    \item \textmd{'g2.v2g\_map'} : $g^2_{\mathrm{map}}$
    \item \textmd{'g2.ilist'} : $i^{(0)}_{\mathrm{+,GSL}}$
    \item \textmd{'g2.jlist'} : $j^{(0)}_{\mathrm{+,GSL}}$
    \item \textmd{'g2.shift'} : $\mathbf{n_+}$
\end{itemize}
$\nijmax$ should be provided in advance. \textit{interactions} is a list of 
\textit{string} representing the \textbf{ordered} interactions (e.g. 
\textmd{['NiNi','NiMo','MoMo','MoNi']}). \textit{offsets} is a list of integers 
marking the starting indices of the \textit{interactions}. As an example, if 
$N_{\eta}$ is 8, the \textit{offsets} corresponding to 
\textmd{['NiNi','NiMo','MoMo','MoNi']} should be \textit{[0, 8, 16, 24]}.

\begin{verbatim}
def get_g2_map(atoms: Atoms,
               rc: float,
               nij_max: int,
               interactions: list,
               vap: VirtualAtomMap,
               offsets: np.ndarray,
               for_prediction=False):
    if for_prediction:
        iaxis = 0
    else:
        iaxis = 1
    g2_map = np.zeros((nij_max, iaxis + 2), dtype=np.int32)
    tlist = np.zeros(nij_max, dtype=np.int32)
    symbols = atoms.get_chemical_symbols()
    tic = time.time()
    ilist, jlist, n1 = neighbor_list('ijS', atoms, rc)
    nij = len(ilist)
    tlist.fill(0)
    for i in range(nij):
        symboli = symbols[ilist[i]]
        symbolj = symbols[jlist[i]]
        tlist[i] = interactions.index('{}{}'.format(symboli, symbolj))
    ilist = np.pad(ilist + 1, (0, nij_max - nij), 'constant')
    jlist = np.pad(jlist + 1, (0, nij_max - nij), 'constant')
    n1 = np.pad(n1, ((0, nij_max - nij), (0, 0)), 'constant')
    n1 = n1.astype(np.float32)
    for count in range(len(ilist)):
        if ilist[count] == 0:
            break
        ilist[count] = vap.index_map[ilist[count]]
        jlist[count] = vap.index_map[jlist[count]]
    g2_map[:, iaxis + 0] = ilist
    g2_map[:, iaxis + 1] = offsets[tlist]
    return {"g2.v2g_map": g2_map, "g2.ilist": ilist, "g2.jlist": jlist,
            "g2.shift": n1}
\end{verbatim}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section Appendix. F
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Implementation of the \texorpdfstring{$g^4_{\mathrm{map}}$}{g4map}}

The construction of $g^4_{\mathrm{map}}$ is similar to $g^2_{\mathrm{map}}$. 
$\nijkmax$ should be provided in advance.

One must note that \textmd{'g2.ilist'} contains GSL indices while the symbol 
list of the target \textit{Atoms} is in local order. So the \textit{reverse} 
mapping is necessary.

\begin{verbatim}
def get_g4_map(atoms: Atoms,
               g2_map: dict,
               interactions: list,
               offsets: np.ndarray,
               vap: VirtualAtomMap,
               nijk_max: int,
               for_prediction=True):
    if for_prediction:
        iaxis = 0
    else:
        iaxis = 1
    g4_map = np.zeros((nijk_max, iaxis + 2), dtype=np.int32)
    ijk = np.zeros((nijk_max, 3), dtype=np.int32)
    n1 = np.zeros((nijk_max, 3), dtype=np.float32)
    n2 = np.zeros((nijk_max, 3), dtype=np.float32)
    n3 = np.zeros((nijk_max, 3), dtype=np.float32)
    symbols = atoms.get_chemical_symbols()
    indices = {}
    vectors = {}
    for i, atom_gsl_i in enumerate(g2_map["g2.ilist"]):
        if atom_gsl_i == 0:
            break
        if atom_gsl_i not in indices:
            indices[atom_gsl_i] = []
            vectors[atom_gsl_i] = []
        indices[atom_gsl_i].append(g2_map["g2.jlist"][i])
        vectors[atom_gsl_i].append(g2_map["g2.shift"][i])
    count = 0
    for atom_gsl_i, nl in indices.items():
        atom_local_i = vap.reverse_map[atom_gsl_i]
        symboli = symbols[atom_local_i]
        prefix = '{}'.format(symboli)
        for j in range(len(nl)):
            atom_vap_j = nl[j]
            atom_local_j = vap.reverse_map[atom_vap_j]
            symbolj = symbols[atom_local_j]
            for k in range(j + 1, len(nl)):
                atom_vap_k = nl[k]
                atom_local_k = vap.reverse_map[atom_vap_k]
                symbolk = symbols[atom_local_k]
                interaction = '{}{}'.format(
                    prefix, ''.join(sorted([symbolj, symbolk])))
                ijk[count] = atom_gsl_i, atom_vap_j, atom_vap_k
                n1[count] = vectors[atom_gsl_i][j]
                n2[count] = vectors[atom_gsl_i][k]
                n3[count] = vectors[atom_gsl_i][k] - vectors[atom_gsl_i][j]
                index = interactions.index(interaction)
                g4_map[count, iaxis + 0] = atom_gsl_i
                g4_map[count, iaxis + 1] = offsets[index]
                count += 1
    return {"g4.v2g_map": g4_map,
            "g4.ij.ilist": ijk[:, 0], "g4.ij.jlist": ijk[:, 1],
            "g4.ik.ilist": ijk[:, 0], "g4.ik.klist": ijk[:, 2],
            "g4.jk.jlist": ijk[:, 1], "g4.jk.klist": ijk[:, 2],
            "g4.shift.ij": n1, "g4.shift.ik": n2, "g4.shift.jk": n3}
\end{verbatim}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section Appendix. G
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Compute the descriptors}

Now we can implements the symmetry function descriptor. Here 
\textit{SymmetryFunction} is the base class and \textit{BatchSymmetryFunction} 
is its subclass for batch training.

The return value of the method \textit{build\_graph} is a dict and its keys are
chemical elements and values are correspond atomic descriptors and atom masks. 
Taking the example of the Ni-Mo dataset with $r_c = 4.6$, batch size 10 and only
10 radial symmetry functions, the corresponding $\nijmax$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$ are 108, 
176 and 7200, respectively. Thus, the return dict should be:
\begin{itemize}
    \item \textmd{'Mo'}: ($\mathbf{G}^{\mathrm{Mo}}$, $\delta^{\mathrm{Mo}}$)
    \item \textmd{'Ni'}: ($\mathbf{G}^{\mathrm{Ni}}$, $\delta^{\mathrm{Ni}}$)
\end{itemize}
where $\mathbf{G}^{\mathrm{Mo}}$ is a float tensor of shape 
\textmd{[10, 176, 40]}, $\delta^{\mathrm{Mo}}$ is a float tensor of shape  
\textmd{[10, 176]}, $\mathbf{G}^{\mathrm{Ni}}$ is a float tensor of shape 
\textmd{[10, 108, 40]} and $\delta^{\mathrm{Ni}}$ is a float tensor of shape  
\textmd{[10, 108]} during the training phase.

The demo script can be found on GitHub: https://github.com/Bismarrck/vap.

\begin{verbatim}
class SymmetryFunction:
    gather_fn = staticmethod(tf.gather)

    def __init__(self, rc, elements, eta=np.array([0.05, 4.0, 20.0, 80.0]),
                 omega=np.array([0.0]), beta=np.array([0.005, ]),
                 gamma=np.array([1.0, -1.0]), zeta=np.array([1.0, 4.0]),
                 angular=True, periodic=True):
        all_kbody_terms, kbody_terms, elements = \
            get_kbody_terms(elements, angular=angular)
        ndim, kbody_sizes = compute_dimension(
            all_kbody_terms, len(eta), len(omega), len(beta), len(gamma),
            len(zeta))
        self.rc = rc
        self.all_kbody_terms = all_kbody_terms
        self.kbody_terms = kbody_terms
        self.elements = elements
        self.n_elements = len(elements)
        self.periodic = periodic
        self.angular = angular
        self.kbody_sizes = kbody_sizes
        self.ndim = ndim
        self.kbody_index = {
            kbody_term: self.all_kbody_terms.index(kbody_term)
            for kbody_term in self.all_kbody_terms}
        self.offsets = np.insert(np.cumsum(kbody_sizes), 0, 0)
        self.radial_indices_grid = ParameterGrid({
            'eta': np.arange(len(eta), dtype=int),
            'omega': np.arange(len(omega), dtype=int)})
        self.angular_indices_grid = ParameterGrid({
            'beta': np.arange(len(beta), dtype=int),
            'gamma': np.arange(len(gamma), dtype=int),
            'zeta': np.arange(len(zeta), dtype=int)})
        self.initial_values = {'eta': eta, 'omega': omega, 'gamma': gamma,
                               'beta': beta, 'zeta': zeta}

    @staticmethod
    def get_pbc_displacements(shift, cells, dtype=tf.float32):
        return tf.matmul(shift, cells, name='displacements')

    def get_rij(self, positions, cells, ilist, jlist, shift, name):
        with tf.name_scope(name):
            dtype = positions.dtype
            Ri = self.gather_fn(positions, ilist, 'Ri')
            Rj = self.gather_fn(positions, jlist, 'Rj')
            Dij = tf.subtract(Rj, Ri, name='Dij')
            if self.periodic:
                pbc = self.get_pbc_displacements(shift, cells, dtype=dtype)
                Dij = tf.add(Dij, pbc, name='pbc')
            with tf.name_scope("safe_norm"):
                eps = tf.constant(1e-8, dtype=dtype, name='eps')
                rij = tf.sqrt(tf.reduce_sum(
                    tf.square(Dij, name='Dij2'), axis=-1) + eps)
                return rij, Dij

    def get_v2g_map(self, features: dict, prefix: str):
        return tf.identity(features[f"{prefix}.v2g_map"], name='v2g_map')

    @staticmethod
    def get_v2g_map_delta(tau):
        return tf.constant([0, tau], dtype=tf.int32, name='delta')

    def get_g_shape(self, features: dict):
        return [features['n_atoms_vap'], self.ndim]

    def get_g2_op_for_tau(self, shape, tau, r, rc2, fc_r, base_v2g_map):
        with tf.name_scope(f"Grid{tau}"):
            grid = self.radial_indices_grid[tau]
            etai = grid['eta']
            omegai = grid['omega']
            eta = tf.constant(self.initial_values['eta'][etai], r.dtype)
            omega = tf.constant(
                self.initial_values['omega'][omegai], r.dtype)
            delta = self.get_v2g_map_delta(tau)
            r2c = tf.math.truediv(tf.square(r - omega), rc2, name='r2c')
            v = tf.exp(-tf.multiply(eta, r2c, 'eta_r2c')) * fc_r
            v2g_map_tau = tf.add(base_v2g_map, delta, f'v2g_map_{tau}')
            return tf.scatter_nd(v2g_map_tau, v, shape, f"g{tau}")

    def get_g2_op(self, features: dict):
        with tf.variable_scope("G2"):
            r = self.get_rij(features['positions'],
                             features['cells'],
                             features['g2.ilist'],
                             features['g2.jlist'],
                             features['g2.shift'],
                             name='rij')[0]
            rc2 = tf.constant(self.rc**2, dtype=r.dtype, name='rc2')
            fc_r = cosine_cutoff(r, rc=self.rc, name='fc_r')
            base_v2g_map = self.get_v2g_map(features, prefix='g2')
            shape = self.get_g_shape(features)
            values = []
            for tau in range(len(self.radial_indices_grid)):
                values.append(
                    self.get_g2_op_for_tau(
                        shape, tau, r, rc2, fc_r, base_v2g_map))
            return tf.add_n(values, name='g')

    def get_g4_op_for_tau(self, shape, tau: int, cos_theta, r2c, fc_r,
                          base_v2g_map):
        with tf.name_scope(f"Grid{tau}"):
            grid = self.angular_indices_grid[tau]
            betai = grid['beta']
            gammai = grid['gamma']
            zetai = grid['zeta']
            beta = tf.constant(
                self.initial_values['beta'][betai], r2c.dtype)
            gamma = tf.constant(
                self.initial_values['gamma'][gammai], r2c.dtype)
            zeta = tf.constant(
                self.initial_values['zeta'][zetai], r2c.dtype)
            delta = self.get_v2g_map_delta(tau)
            one = tf.constant(1.0, dtype=r2c.dtype, name='one')
            two = tf.constant(2.0, dtype=r2c.dtype, name='two')
            gt = tf.math.multiply(gamma, cos_theta, name='gt')
            gt1 = tf.add(gt, one, name='gt1')
            gt1z = tf.pow(gt1, zeta)
            z1 = tf.math.subtract(one, zeta, name='z1')
            z12 = tf.pow(two, z1)
            c = tf.math.multiply(gt1z, z12, name='c')
            v = tf.multiply(c * tf.exp(-beta * r2c), fc_r, f'v_{tau}')
            v2g_map_tau = tf.add(
                base_v2g_map, delta, name=f'v2g_map_{tau}')
            return tf.scatter_nd(v2g_map_tau, v, shape, f'g{tau}')

    def get_g4_op(self, features: dict):
        with tf.variable_scope("G4"):
            rij = self.get_rij(features['positions'],
                               features['cells'],
                               features['g4.ij.ilist'],
                               features['g4.ij.jlist'],
                               features['g4.shift.ij'],
                               name='rij')[0]
            rik = self.get_rij(features['positions'],
                               features['cells'],
                               features['g4.ik.ilist'],
                               features['g4.ik.klist'],
                               features['g4.shift.ik'],
                               name='rik')[0]
            rjk = self.get_rij(features['positions'],
                               features['cells'],
                               features['g4.jk.jlist'],
                               features['g4.jk.klist'],
                               features['g4.shift.jk'],
                               name='rjk')[0]
            rij2 = tf.square(rij, name='rij2')
            rik2 = tf.square(rik, name='rik2')
            rjk2 = tf.square(rjk, name='rjk2')
            rc2 = tf.constant(self.rc**2, dtype=rij.dtype, name='rc2')
            r2 = tf.add_n([rij2, rik2, rjk2], name='r2')
            r2c = tf.math.truediv(r2, rc2, name='r2_rc2')
            with tf.name_scope("CosTheta"):
                upper = tf.subtract(rij2 + rik2, rjk2, name='upper')
                lower = tf.multiply(2.0 * rij, rik, name='lower')
                cos_theta = tf.math.truediv(upper, lower, name='theta')
            with tf.name_scope("Cutoff"):
                fc_rij = cosine_cutoff(rij, rc=self.rc, name='fc_rij')
                fc_rik = cosine_cutoff(rik, rc=self.rc, name='fc_rik')
                fc_rjk = cosine_cutoff(rjk, rc=self.rc, name='fc_rjk')
                fc_r = tf.multiply(fc_rij, fc_rik * fc_rjk, 'fc_r')
            base_v2g_map = self.get_v2g_map(features, prefix='g4')
            shape = self.get_g_shape(features)
            values = []
            for tau in range(len(self.angular_indices_grid)):
                values.append(
                    self.get_g4_op_for_tau(
                        shape, tau, cos_theta, r2c, fc_r, base_v2g_map))
            return tf.add_n(values, name='g')

    def get_row_split_sizes(self, features: dict):
        return features['row_splits']

    @staticmethod
    def get_row_split_axis():
        return 0

    def get_column_split_sizes(self):
        column_splits = {}
        for i, element in enumerate(self.elements):
            column_splits[element] = [len(self.elements), i]
        return column_splits

    @staticmethod
    def get_column_split_axis():
        return 1

    def split_descriptors(self, descriptors, features: dict):
        with tf.name_scope("Split"):
            row_split_sizes = self.get_row_split_sizes(features)
            row_split_axis = self.get_row_split_axis()
            column_split_sizes = self.get_column_split_sizes()
            column_split_axis = self.get_column_split_axis()
            splits = tf.split(
                descriptors, row_split_sizes, axis=row_split_axis,
                name='rows')[1:]
            atom_masks = tf.split(
                features['atom_masks'], row_split_sizes, 
                axis=row_split_axis,
                name='atom_masks')[1:]
            if len(self.elements) > 1:
                blocks = []
                for i in range(len(splits)):
                    element = self.elements[i]
                    size_splits, idx = column_split_sizes[element]
                    block = tf.split(splits[i],
                                     size_splits,
                                     axis=column_split_axis,
                                     name='{}_block'.format(element))[idx]
                    blocks.append(block)
            else:
                blocks = splits
            return dict(zip(self.elements, zip(blocks, atom_masks)))

    def build_graph(self, features: dict):
        with tf.variable_scope("Behler"):
            descriptors = self.get_g2_op(features)
            if self.angular:
                descriptors += self.get_g4_op(features)
        return self.split_descriptors(descriptors, features)

class BatchSymmetryFunction(SymmetryFunction):
    gather_fn = staticmethod(tf.batch_gather)

    def __init__(self, rc, max_occurs: Counter, nij_max: int, 
                 nijk_max: int, batch_size: int, 
                 eta=np.array([0.05, 4.0, 20.0, 80.0]),
                 omega=np.array([0.0]), beta=np.array([0.005, ]),
                 gamma=np.array([1.0, -1.0]), zeta=np.array([1.0, 4.0]),
                 angular=True, periodic=True):
        elements = sorted(list(max_occurs.keys()))

        super(BatchSymmetryFunction, self).__init__(
            rc=rc, elements=elements, eta=eta, beta=beta, gamma=gamma,
            zeta=zeta, omega=omega, angular=angular, periodic=periodic)
        self._max_occurs = max_occurs
        self._max_n_atoms = sum(max_occurs.values())
        self._nij_max = nij_max
        self._nijk_max = nijk_max
        self._batch_size = batch_size

    @staticmethod
    def get_pbc_displacements(shift, cells, dtype=tf.float32):
        with tf.name_scope("Einsum"):
            shift = tf.convert_to_tensor(shift, dtype=dtype, name='shift')
            cells = tf.convert_to_tensor(cells, dtype=dtype, name='cells')
            return tf.einsum(
                'ijk,ikl->ijl', shift, cells, name='displacements')

    def get_g_shape(self, _):
        n_atoms_vap = self._max_n_atoms + 1
        return [self._batch_size, n_atoms_vap, self.ndim]

    def get_v2g_map_batch_indexing_matrix(self, prefix='g2'):
        if prefix == 'g2':
            ndim = self._nij_max
        else:
            ndim = self._nijk_max
        indexing_matrix = np.zeros(
            (self._batch_size, ndim, 3), dtype=np.int32)
        for i in range(self._batch_size):
            indexing_matrix[i] += [i, 0, 0]
        return indexing_matrix

    @staticmethod
    def get_v2g_map_delta(tau):
        return tf.constant([0, 0, tau], tf.int32, name='delta')

    def get_v2g_map(self, features: dict, prefix="g2"):
        indexing = self.get_v2g_map_batch_indexing_matrix(prefix=prefix)
        return tf.add(
            features[f"{prefix}.v2g_map"], indexing, name='v2g_map')

    def get_row_split_sizes(self, _):
        row_splits = [1, ]
        for i, element in enumerate(self.elements):
            row_splits.append(self._max_occurs[element])
        return row_splits

    @staticmethod
    def get_row_split_axis():
        return 1

    @staticmethod
    def get_column_split_axis():
        return 2
\end{verbatim}

\end{document}