### 2019-10-18 doc/papers/ml-eam/manuscript.tex
The ML-EAM manuscript is updated. The first version is finally finished.

\documentclass[prb,reprint]{revtex4-2}

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
\raggedbottom           % don't add extra vertical space
\begin{comment}
\end{comment}

\bibliographystyle{apsrev4-2}

\begin{document}

\title{
    Machine-learned embedded atom method
}

\author{Xin Chen, 
        Li-Fang Wang, 
        De-Ye Lin$^\mathrm{*}$, 
        Hai-Feng Song$^\mathrm{*}$}

\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}

\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
\end{abstract}

\maketitle

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 1. Introduction
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Introduction}
\label{sec:introduction}

Atomistic modeling plays a vital role in materials science. \textit{ab initio} 
calculation or force-field based molecular dynamics simulation (MD) are 
effective ways to study, understand or predict chemical and physical properties 
of materials. \textit{ab initio} approaches are generally much more precise but 
they are rarely used on large-scale metallic systems due to their extremely-high 
computation expenses. Physical model based empirical potential (force-field), 
such as the embedded-atom method (EAM), modified embedded-atom method (MEAM), 
bond-order potential (BoP), or angular-dependent potential (ADP), still plays 
the major role in long-time simulations and these empirical methods can achieve 
reasonable accuracy with much lower computation costs. Empirical potentials 
generally have very few learnable parameters and both microscopic observables 
(energy, forces, virial, etc.) and macroscopic observables (melting 
point, surface energy, etc.) can be used to tune these parameters. Finding 
optimal parameters of empirical potentials is always a challenging task. Global 
optimization (GO) approaches (Basin-Hopping, genetic algorithm, etc) are 
traditionally used to find the best possible parameters. However, the gradients 
of the losses with respect to model parameters are difficult or even impossible 
to calcualte. Hence, GO optimizations are generally not that effective.

In the last decade, machine learning (ML) has become one of the hottest 
topics in many research areas. In the materials science, researchers have made 
great efforts on developing ML models to describe atomic interactions. Such ML 
models are considered as machine learning interaction potentials (MLIPs). Until 
now, hundreds of MLIPs have been proposed. Among them, the symmetry-function 
based atomistic neural network (ANN) model, published by Parinello and Behler in 
2007, is still the most popular choice in modeling metallic interactions. 
The smooth overlap atomic positions descriptor based gaussian approximation 
potential (SOAP-GAP), developed by Bartòk et al, can give extremely accurate 
prediction results, although it's a bit computationally expensive. 
Recently, Thompson and co-workers proposed another quantum-accurate MLIP named 
the spectral neighbor analysis potential (SNAP) and it has been proven working  
on a broad range of metals and alloys.

In many cases, MILPs can easily outperform state-of-art empirical potentials. 
Compared with empirical potentials, MLIPs generally have orders of magnitudes 
more model paramters. The redundant parameter space greatly reduces the 
difficulty of fitting complicated potential energy surfaces. But, to effectively 
train a MLIP and avoid overfitting, a large high-quality (versatile) training 
dataset is probably needed. However, MILPs \textbf{can really} take advantages 
of "big data" for two reasons. First, MLIPs typically only have 
\textit{basic} or \textit{simple} arithmetic operations. Thus, MILPs can be 
implemented within modern deep learning frameworks (TensorFlow, PyTorch, etc) 
so that the gradients of the total loss with respect to fitting parameters can 
be can obtained with the backpropagation algorithm automatically and 
efficiently. Second, MLIPs are mostly vectorizable. Hence, GPUs can be utilized
to significantly accelerate training and using of MILPs.

However, MLIPs also have challenges. The large parameter space and lack of 
physical background makes the "big data" a necessarity. The cost of dataset is 
non-negligible. Besides, even "big data" can only cover a small portion of real 
physical environments (temperature, external pressure, etc). Outside the 
training zone, the performances of MILPs may not that stable. For long-time 
molecular dynamics (MD) simulations of large-scale ($10^5$ or more) systems, 
computation efficiency also becomes a major concern. Recent benchmark tests 
suggest that MILPs are still too expensive. At present, most MLIPs are used to 
examine small to medium ($10^3$ to $10^4$) systems.

In this work, instead of designing new atomic descriptors, we chose a new route 
to develop MLIP: combining machine learning with empirical potentials. We 
successfully implemented EAM and its variant ADP within TensorFlow so that 
machine learning approaches can be used directly to tune EAM and ADP potentials. 
Our results suggest ML-EAM or ML-ADP can be as precise as the SNAP machine 
learning method.

This paper is organized as follows. Section \ref{sec:method} describes the 
theoretical background of this work, including the formalism of the embedded 
atom method and algorithms and details of the machine learned EAM. 
Section \ref{sec:results} summarizes the training results and the optimal 
parameters. Applications of the new potentials are discussed in 
Section \ref{sec:discussions}.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 2. Method
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Method}
\label{sec:method}

% ------------------
% Section 2.A Theory
% ------------------
\subsection{Theory}
\label{sec:theory}

In the original EAM formalism, the total energy, $E^{total}$, is the sum of 
atomic energies:
\begin{align}
E^{total} = & \sum_{i}^{N}{E_{i}} \nonumber \\
\label{eq:eam_total_energy}
= & \sum_{i}^{N}{F_{a}(\rho_i)} + 
    \frac{1}{2}\sum_{i}^{N}{\sum_{j \neq i}^{r_{ij} < r_c}{
    \phi_{ab}(r_{ij})
}}
\end{align}
where $r_c$ is the cutoff radius, $a$ and $b$ represents species of atoms $i$ 
and $j$, $\phi_{ab}(r_{ij})$ is energy of the pairwise interaction between $i$ 
and $j$, $F_{a}(\rho_{i})$ is the embedding energy and $\rho_{i}$ is the local 
electron density of atom $i$. $\rho_{i}$ can be calculated with the following 
equation:
\begin{equation}
\label{eq:rho_eam}
\rho_{i} = \sum_{j}^{r_{ij} < r_{c}}{
    \rho_{b}(r_{ij})
}
\end{equation}
where $\rho_{b}$ is the electron density function of specie $b$. In the 
Finnis-Sinclair model, the electron density has a slightly modified form:
\begin{equation}
\label{eq:rho_fs}
\rho_{i}^{\mathrm{FS}} = \sum_{j}^{r_{ij} < r_{c}}{
    \rho_{ab}(r_{ij})
}
\end{equation}
$F$, $\rho$ and $\phi$ can be either parameterized functions or cubic splines.

The original EAM formalism does not inlcude angular-dependent interactions. To 
fix this problem, Baskes modified the original EAM and got MEAM (modified 
embedded-atom method), Lenosky proposed an alternative spline-based interpretion 
of MEAM  while Mishin developed the angular-dependent potential (ADP). The ADP 
formalism introduces three additional terms to the total energy:
\begin{align}
\label{eq:adp}
E^{total} 
& = E^{\mathrm{EAM}} \nonumber \\
& + \frac{1}{2}\sum_{i}{\sum_{\alpha}{(\mu_{i}^{\alpha})^2}} \nonumber \\
& + \frac{1}{2}\sum_{i}{
    \sum_{\alpha}{\sum_{\beta}{(\lambda_{i}^{\alpha\beta})^2}}} \nonumber \\
& - \frac{1}{6}\sum_{i}{\nu_{i}^{2}}
\end{align}
These terms represent non-central bonding contributions and they can be computed
with the following equations:
\begin{align}
\label{eq:adp_mu}
\mu_{i}^{\alpha} & = \sum_{j \neq i}{\mu_{ab}(r_{ij}) r_{ij}^{\alpha}} \\
\label{eq:adp_lambda}
\lambda_{i}^{\alpha\beta} & = \sum_{j \neq i}{
    \omega_{ab}(r_{ij}) r_{ij}^{\alpha}r_{ij}^{\beta}} \\
\label{eq:adp_nu}
\nu_{i} & = \sum_{\alpha}{\lambda_{i}^{\alpha\alpha}}
\end{align}
where $\mu_{ab}(r)$ and $\omega_{ab}(r)$ can be viewed as measures of the 
strengths of dipole and quadrupole interactions.

% -----------
% Section 2.B Transformation
% -----------
\subsection{Transformation}
\label{sec:transformation}

To integrate EAM/ADP with machine learning, the original total energy expression 
(Equation \ref{eq:eam_total_energy}) must be transformed to a vectorizable form. 
Without loss of generality, we take the binary alloy, AB, to demonstrate how to 
do the transformation.

\newcommand{\niaa}{N_{i}^{\mathrm{AA}}}
\newcommand{\niab}{N_{i}^{\mathrm{AB}}}
\newcommand{\njbb}{N_{j}^{\mathrm{BB}}}
\newcommand{\njba}{N_{j}^{\mathrm{BA}}}
\newcommand{\riaa}{\vec{\mathbf{r}}_{i}^{\mathrm{AA}}}
\newcommand{\riab}{\vec{\mathbf{r}}_{i}^{\mathrm{AB}}}
\newcommand{\ribb}{\vec{\mathbf{r}}_{i}^{\mathrm{BB}}}
\newcommand{\riba}{\vec{\mathbf{r}}_{i}^{\mathrm{BA}}}
\newcommand{\namax}{N_{\mathrm{A}}^{\mathrm{max}}}
\newcommand{\nbmax}{N_{\mathrm{B}}^{\mathrm{max}}}
\newcommand{\nnl}{N^{\mathrm{nl}}}
\newcommand{\nnli}{N^{\mathrm{nl}}_i}

Suppose the cutoff radius $r_c$ is fixed, the energy of atom $i$ of specie $A$ 
can be calculated with the following expanded equation:
\begin{align}
\label{eq:e_atom_phase_1}
E_{i}^{A} 
= & 
\frac{1}{2}\sum_{j\neq i}^{\niaa}\phi_{\mathrm{AA}}(r_{ij}) +
\frac{1}{2}\sum_{j\neq i}^{\niab}\phi_{\mathrm{AB}}(r_{ij}) \nonumber \\
+ & F_{\mathrm{A}}\left(
    \sum_{j\neq i}^{\niaa}{\rho_{\mathrm{A}}(r_{ij})} +
    \sum_{j\neq i}^{\niab}{\rho_{\mathrm{B}}(r_{ij})} 
\right)
\end{align}
where $\niaa$ represents the number of A-type neighors of atom $i$ and 
$\niab$ represents the number of B-type neighors. For atom $j$ of specie $B$, we
can also write a similar form. 
\begin{align}
\label{eq:e_atom_phase_1}
E_{j}^{B} 
= & 
\frac{1}{2}\sum_{i \neq j}^{\njbb}\phi_{\mathrm{BB}}(r_{ij}) +
\frac{1}{2}\sum_{i \neq j}^{\njba}\phi_{\mathrm{AB}}(r_{ij}) \nonumber \\
+ & F_{\mathrm{B}}\left(
    \sum_{j\neq i}^{\njbb}{\rho_{\mathrm{B}}(r_{ij})} +
    \sum_{j\neq i}^{\njba}{\rho_{\mathrm{A}}(r_{ij})} 
\right)
\end{align}
When $r_c$ is fixed, $\niaa$, $\niab$, $\njbb$ and $\njba$ are all constants and
$\nnli$ is the maximum of these numbers. Finally, we can pre-determine the 
maximum neighor list size $\nnl$:
\begin{equation}
\label{eq:nnl}
\nnl = \max{\left(\nnli\right)}
\end{equation}
$\nnl$ is also a constant in the training phase because both the training 
dataset and $r_c$ are fixed.

Next, assume $H(x)$ represents the heaviside step function:
\begin{equation}
\label{eq:heaviside}
H(x) = \begin{cases}
    1 & x > 0 \\
    0 & x \leq 0 \\
\end{cases}
\end{equation}
then the pairwise term can be transformed to:
\begin{align}
\sum_{j\neq i}^{\niaa}\phi_{\mathrm{AA}}(r_{ij}) = &
\sum_{j\neq i}^{\niaa}\phi_{\mathrm{AA}}(r_{ij}) \cdot 1 + 
\sum_{j\neq i}^{\nnl - \niaa}\phi_{\mathrm{AA}}(0) \cdot 0 \nonumber \\
= & \phi_{\mathrm{AA}}(\vec{\mathbf{r}}_{i}^{\mathrm{AA}})^T
H(\vec{\mathbf{r}}_{i}^{\mathrm{AA}})
\end{align}
where $\vec{\mathbf{r}}_{i}^{\mathrm{AA}}$ is a $\nnl$-length column vector 
whose last $\nnl - \niaa$ elements are zeros.

We can write Equation \ref{eq:e_atom_phase_1} in an equivalent expression:
\begin{align}
E_{i}^{\mathrm{A}} 
\label{eq:e_atom_A}
= &
\frac{1}{2}\left( 
    \phi_{\mathrm{AA}}(\vec{\mathbf{r}}_{i}^{\mathrm{AA}})^T
    H(\vec{\mathbf{r}}_{i}^{\mathrm{AA}}) +
    \phi_{\mathrm{AB}}(\vec{\mathbf{r}}_{i}^{\mathrm{AB}})^T
    H(\vec{\mathbf{r}}_{i}^{\mathrm{AB}})
\right) \nonumber \\
+ &
F_{\mathrm{A}}\left( 
    \rho_{\mathrm{AA}}(\vec{\mathbf{r}}_{i}^{\mathrm{AA}})^T
    H(\vec{\mathbf{r}}_{i}^{\mathrm{AA}}) +
    \rho_{\mathrm{AB}}(\vec{\mathbf{r}}_{i}^{\mathrm{AB}})^T
    H(\vec{\mathbf{r}}_{i}^{\mathrm{AB}})
\right)
\end{align}
Here $\vec{\mathbf{r}}_{i}^{\mathrm{AB}}$ is also a $\nnl$-length vector. For 
atom $j$ of specie $B$, we can also derive its energy $E^{\mathrm{B}}_{j}$:
\begin{align}
E_{j}^{\mathrm{B}}
\label{eq:e_atom_B}
= &
\frac{1}{2}\left( 
    \phi_{\mathrm{BB}}(\vec{\mathbf{r}}_{j}^{\mathrm{BB}})^T
    H(\vec{\mathbf{r}}_{j}^{\mathrm{BB}}) +
    \phi_{\mathrm{BA}}(\vec{\mathbf{r}}_{j}^{\mathrm{BA}})^T
    H(\vec{\mathbf{r}}_{j}^{\mathrm{BA}})
\right) \nonumber \\
+ &
F_{\mathrm{A}}\left( 
    \rho_{\mathrm{BB}}(\vec{\mathbf{r}}_{j}^{\mathrm{BB}})^T
    H(\vec{\mathbf{r}}_{j}^{\mathrm{BB}}) +
    \rho_{\mathrm{BA}}(\vec{\mathbf{r}}_{j}^{\mathrm{BA}})^T
    H(\vec{\mathbf{r}}_{j}^{\mathrm{BA}})
\right)
\end{align}

Since $\riaa$, $\riab$, $\ribb$ and $\riba$ all have the same length ($\nnl$), 
we can use a (redundant) matrix, $\mathbf{g}_{i}$, to describe all neighbors of 
atom $i$:
\begin{equation}
\mathbf{g}_i = \begin{bmatrix}
    \riaa & \riab & \ribb & \riba \\
\end{bmatrix}
\end{equation}
$\mathbf{g}_i$ is a $\nnl \times 4$ matrix. If the specie of atom $i$ is A, only
the first two columns have non-zero values. Similarly, the last two columns will
have non-zeros values if atom $i$ is a B-type atom. In fact, $\mathbf{g}_i$ can
be viewed as the EAM descriptors for atom $i$. Hence, each structure can be 
expressed with a 3D matrix, $\mathbf{G}$ , of shape $N \times \nnl \times 4$.

During the training phase, the maximum appearances of element A and B in any 
structure ($\namax$ and $\nbmax$) are also constants. Thus, any $\mathbf{G}$ can 
be expanded to a $(\namax + \nbmax) \times \nnl \times 4$ matrix $\mathbf{G}'$ 
by zero-padding. In summary, arbitrary structure in the training dataset can be 
converted to a fixed-shape descriptor matrix $\mathbf{G}'$.

For the ADP formalism, the corresponding transformation is almost the same 
except that $\mathbf{g}_{i}^{\mathrm{adp}}$ should also include the XYZ 
components of $\vec{\mathbf{r}}_i$.

% -----------
% Section 2.C Functions
% -----------
\subsection{Functions}
\label{sec:functions}

In this work, we use the EAM potential published by Zhou, Johnson and Wadley 
(Zjw04) as an example to validate our machine learning approach. Zjw04 is a 
quite popular EAM potential. In the Zjw04 potential, the electron density 
function has the following form:
\begin{equation}
\label{eq:zjw04_rho}
\rho_{b}(r) = \frac{
    f_{e} \exp\left[ -\beta\left( r/r_{e} - 1 \right) \right]
}{
    1 + \left(r/r_{e} - \lambda\right)^{20}
}
\end{equation}
where $r_e$ is a constant equal to equilibrium spacing between nearest neighors, 
$f_{e}$, $\beta$, $\lambda$ are adjustable parameters. The pairwise potential 
between the same species can be computed with:
\begin{equation}
\label{eq:zjw04_phi_aa}
\phi_{aa}(r) = 
\frac{A \exp\left[ -\alpha\left( r/r_{e} - 1 \right) \right]}
     {1 + \left(r / r_{e} - \kappa\right)^{20}} - 
\frac{B \exp\left[ -\beta\left( r/r_{e} - 1 \right) \right]}
     {1 + \left(r / r_{e} - \lambda\right)^{20}}
\end{equation}
where $A$, $B$, $\alpha$ and $\kappa$ are also trainable parameters, $\beta$ and 
$\kappa$ are used in Equation \ref{eq:zjw04_rho} already. For the pairwise 
interaction between two atoms of different species, Zhou et al proposed an 
interpolation form:
\begin{equation}
\label{eq:zjw04_phi_ab}
\phi_{ab}(r) = \frac{1}{2}\left(
    \frac{\rho_{b}(r)}{\rho_{a}(r)}\phi_{aa}(r) +
    \frac{\rho_{a}(r)}{\rho_{b}(r)}\phi_{bb}(r) 
\right)
\end{equation}
The embedding function has a more complicated form as it requires to fit a much 
wider range of electron density values:
\begin{equation}
\label{eq:zjw04_embed}
F(\rho) = \begin{cases}
    \sum_{i=0}^{3}{F_{ni}\left( \frac{\rho}{\rho_n} - 1 \right)^{i}} &
    \rho < \rho_{n} \\
    \sum_{i=0}^{3}{F_{i}\left( \frac{\rho}{\rho_e} - 1 \right)^{i}} &
    \rho_{n} \leq \rho < \rho_{0} \\
    F_{e}\left[ 
        1 - \eta\ln\left( \frac{\rho}{\rho_s}\right) 
    \right](\frac{\rho}{\rho_s})^{\eta} & \rho_{0} \leq \rho
\end{cases}
\end{equation}
where $F_{ni}$, $F_{i}$, $\rho_e$, $\rho_s$, $\eta$ and $F_{e}$ are trainable 
parameters, $\rho_n=0.85\rho_e$ and $\rho_0=1.15\rho_e$. For each metal, there 
are 15 adjustable parameters. The original embedding potential is a stepwise 
function. Thus, the minimization requires some tricks to ensure its continuity.
To make it simpler, we slightly modified Equation \ref{eq:zjw04_embed}:
\begin{align}
\label{eq:zjw04xc_embed}
F(\rho) 
= & c_1 \cdot
\sum_{i=0}^{3}{F_{ni}\left( \frac{\rho}{\rho_n} - 1 \right)^{i}} \nonumber \\
+ & c_2 \cdot 
\sum_{i=0}^{3}{F_{i}\left( \frac{\rho}{\rho_e} - 1 \right)^{i}} \nonumber \\
+ & c_3 \cdot
F_{e}\left[1 - \eta\ln\left( \frac{\rho}{\rho_s}\right)\right]
(\frac{\rho}{\rho_s})^{\eta} \\
\label{eq:frho_c1}
c_{1} = & \frac{1}{1 + e^{-2\left(\rho_{n} - \rho\right)}} \\
\label{eq:frho_c3}
c_{3} = & \frac{1}{1 + e^{-2\left(\rho - \rho_{0}\right)}} \\
c_{2} = & 1 - c_1 - c_3
\end{align}
$\omega_1$ and $\omega_3$ are just damping factors calculated by the sigmoid 
functions (Equations \ref{eq:frho_c1} and \ref{eq:frho_c3}).

The dipole ($\mu_{ab}$) and quadrupole ($\lambda_{ab}$) functions have the same 
form (developed by Mishin):
\begin{align}
\label{eq:adp_dipole}
\mu_{ab}(r) & = \left[
    d_{1}^{ab} \exp\left( -d_{2}^{ab}r \right) + d_{3}^{ab}
\right] \psi\left( \frac{r - r_{0}}{r_{h}} \right) \\
\omega_{ab}(r) & = \left[
    q_{1}^{ab} \exp\left( -q_{2}^{ab}r \right) + q_{3}^{ab}
\right] \psi\left( \frac{r - r_{0}}{r_{h}} \right)
\end{align}
where $d_{i}$, $q_{i}$, $r_0$ and $r_{h}$ are trainable parameters and $\psi(x)$
is a damping function:
\begin{equation}
\label{eq:mishin_cutoff}
\psi(x) = \begin{cases}
    0 & x \ge 0 \\
    \frac{x^4}{1 + x^4} & x < 0 
\end{cases}
\end{equation}

% -----------
% Section 2.D Physical constraints
% -----------
\subsection{Physical constraints}
\label{sec:constraints}

Physical constraints are quite common in fitting traditional empirical 
potentials. Physical constraints are typically static (cohesive energy, elastic 
constants, etc) collected from experiments and they can be very effective when 
training data is limited. For example, the cohesive energy, bulk modulus, 
vacancy formation energy and other constraints were used to develop the original 
Zjw04 potential. Mishin et al adopted the Rose universal equation of state to 
ensure the performaces of his ADP potentials in the high pressure region. 
However, such constraints are really rare in developing MILPs. One possible 
explanation may be these constraints are generally derived properties and 
implementing them in the loss function are technically difficult.

In this work, we successfully integrated two constraints into the total loss:
the Rose equation of state (EOS) constraint and the elastic tensor constraint. 
These two losses will be discussed later. The details of their implementations 
will be described in another paper.

The Rose constraint incoporates the universal equation of state (Rose et al) 
into the total loss function. The Mishin-modified equation 
\begin{equation}
\label{eq:rose_eos}
E(x) = E_{0}\left[
    1 + \alpha x + \beta \alpha^3 x^3 \frac{2x + 3}{(x - 1)^2} \right]
    e^{-\alpha x}
\end{equation}
is used because the original form tends to underestimate energies under high 
pressures. In Equation \ref{eq:rose_eos}, $E_{0}$ is the energy of the 
equilibrium structure, $x = a / a_{0} - 1$ is the relative isotropic scaling 
factor ($a$ is the lattice constant), $\beta$ is a chosen parameter and 
\begin{equation}
\label{eq:rose_alpha}
\alpha = \sqrt{-\frac{9 V_{0} B }{E_{0}}}
\end{equation}
where $V_0$ is the equilibrium volume and $B$ is the bulk modulus. The adoption
of the Rose constraint guarantees the exact predictions of bulk modulus and 
the energy-volume curve. The loss of the Rose EOS constraint 
$\mathbf{L}^{\mathrm{Rose}}$  is measured as the 2-norm of the energy 
differences between $E(x)^{\mathrm{dft}}$ and their corresponding predicted 
$E(x)$: 
\begin{equation}
\label{eq:rose_loss}
\mathbf{L}^{\mathrm{Rose}} = \sum_{\Omega}{
    \sqrt{\sum_{i}{\left(E(x_i) - E(x_i)^{\mathrm{dft}}\right)^2}}
}
\end{equation}
Here $\Omega$ denotes all selected crystals. In this work, for each included 
crystal, we use the same choices of $x$: 
$x_{i} = x_{0} + \Delta x \cdot i, x_{0} = -0.1, N_{t}^{\mathrm{max}} = 20, 
\Delta x=0.01$. In this work, $\beta^{\mathrm{Ni}}$ is 0.005 and 
$\beta^{\mathrm{Mo}}$ is 0.150.

Elastic tensor is also a popular constraint for tuning empirical potentials but 
rarely used directly in optimizing MILPs. Shyue Ping Ong used this constraint in 
the outer loop (the global optimization based property-matching step) to find 
optimal parameters of SNAP potentials.

In this work, we successfully find a way to use elastic tensor directly as a 
constraint. Given an equilibrium crystal structure, its elastic constant 
$c_{ijkl}$ can be derived from $E^{total}$ directly:
\begin{align}
\label{eq:virial}
V \cdot \epsilon & = -\mathbf{F}^{T}\mathbf{R} + \left(
    \frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h} \\ 
\label{eq:cijkl}
c_{ijkl} |_{\epsilon \to 0, \mathbf{F} \to 0} & = \frac{1}{V}\left[
    \left( 
        \frac{\partial{\epsilon_{ij}}}{\partial{\mathbf{h}}}
    \right)^{\mathrm{T}}\mathbf{h}
\right]_{kl}
\end{align}
where $V$ is the volume, $\epsilon$ is the $3 \times 3$ virial stress tensor, 
$\mathbf{h}$ is the \textit{row-major} $3 \times 3$ lattice tensor, $\mathbf{F}$ 
and $\mathbf{R}$ are $N \times 3$ matrices representing the total forces and 
atomic positions.

In this work, the loss $\mathbf{L}^{\mathrm{elastic}}$ contributed by the 
elastic tensor is also measured by the RMSE between $c_{ijkl}$ and 
$c_{ijkl}^{\mathrm{dft}}$:
\begin{align}
\label{eq:cijkl_loss}
\mathbf{L}^{\mathrm{elastic}} 
& = \sum_{\Omega}{\left(
    \omega_{\Omega} \cdot \mathbf{RMSE}_{\Omega} 
    + ||\epsilon|| + ||\mathbf{F}||
\right)}
 \\
\label{eq:cijkl_loss_gate}
\omega_{\Omega} & = \mathbf{ReLU}(\mathbf{MAE}_{\Omega} - \tau) \\
\label{eq:relu}
\mathbf{ReLU}(x) & = \begin{cases}
    x & x \ge 0 \\
    0 & x < 0
\end{cases}
\end{align}
where $\sum_{\Omega}$ loops through all included crystals, 
$\mathbf{MAE}_{\Omega}$ and $\mathbf{RMSE}_{\Omega}$ are the mean absolute error 
and root mean squared error between predicted elastic constants and DFT elastic 
constants. $\tau$ is a pre-selected gate parameter. 
When $\mathbf{MAE}$ is below $\tau$, $\mathbf{L}^{\mathrm{elastic}}$ will not 
contribute to the total loss. In this work, $\tau$ is set to 2 GPa.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EAM - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.57]{figures/Ni_eam.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni - Material Properties
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}
\centering
\begin{tabular}{lccccc}
\hline
                              & DFT  & SNAP          & EAM            & ML-EAM         & Exp.      \\
\hline
$T_{m}$(K)                    &      & 1785          & 1520           & 1520           & 1728      \\
$c_{11}$ (GPa)                & 276  & 276 (0.0\%)   & 248 (-10.1\%)  & 274 (-0.7\%)   & 261       \\
$c_{12}$ (GPa)                & 159  & 159 (0.0\%)   & 147 (-7.5\%)   & 163 (2.5\%)    & 151       \\
$c_{44}$ (GPa)                & 132  & 132 (0.0\%)   & 125 (-5.3\%)   & 131 (-0.8\%)   & 132       \\
$B_{\mathrm{VRH}}$ (GPa)      & 198  & 198 (0.0\%)   & 181 (-8.6\%)   & 195 (0.0\%)    & 188       \\
$G_{\mathrm{VRH}}$ (GPa)      & 95   & 95 (0.0\%)    & 87 (-8.4\%)    & 93 (-2.1\%)    & 93        \\
$\mu$                         & 0.29 & 0.29 (0.0\%)  & 0.29 (0.0\%)   & 0.29 (0.0\%)   & 0.29      \\
$E_{v}$ (eV)                  & 1.46 & 1.68 (15.1\%) & 1.68 (15.1\%)  & 1.71 (17.1\%)  & 1.54–1.80 \\
$E_{m}$ (eV)                  & 1.12 & 1.07 (-4.5\%) & 0.90 (-19.6\%) & 0.87 (-22.3\%) & 1.01–1.48 \\
$E_{a} = E_{v} + E_{m} $ (eV) & 2.58 & 2.75 (6.6\%)  & 2.58 (0.0\%)   & 2.58 (0.0\%)   & 2.77–2.95 \\
\hline
\end{tabular}
\caption{\label{table:Ni_properties} Comparison of the calculated and 
experimental material properties of elementary fcc Ni. $T_{m}$ is the melting 
point. $c_{ij}$ represents elastic constants. $E_{v}$ and $E_{m}$ are vacancy 
formation energy and migration energy, respectively. SNAP refers to the 
elementary Ni SNAP model.
}
\end{table*}

% -----------
% Section 2.E Implementation
% -----------
\subsection{Implementation}
\label{sec:implementation}

The implementation of the machine learned embedded atom mathod and the physical
constraints are beyond the scope of this work. The details will be discussed
in another paper. Here, we will give a brief description.

Both ML-EAM and ML-ADP are implemented within Google's TensorFlow. The 
virtual-atom approach (VAP) is adopted so that we can construct a computation 
graph from atomic positions to total energy directly. Thus, atomic forces 
(Equation \ref{eq:forces}) and virial stress (Equation \ref{eq:virial}) can be 
derived by the AutoGrad module of TensorFlow and calcualted automatically and 
efficiently. 
\begin{equation}
\label{eq:forces}
\mathbf{F} = -\frac{\partial E^{total}}{\partial \mathbf{R}}
\end{equation}
This direct computation graph also plays a key role in computing analytical 
elastic tensor (Equation \ref{eq:cijkl}) and the Rose loss 
(Equation \ref{eq:rose_eos}).

To find optimal parameters of the funcitons in section \ref{sec:functions}, the 
overall loss function shall be defined. Just like other machine learning tasks, 
the mini-batch stochastic gradient descent algorithm is used to minimize the 
loss function. Equation \ref{eq:loss} demonstrates the loss function used in 
this work:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3\sum_{i}^{N_{b}}{N_i}}\sum_{i}^{N_b}{\sum_{j}^{N_i}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{s}}\sqrt{\frac{1}{6N_b}\sum_{i}^{N_b}{
    \sum_{j}^{6}{
        \left(
            \epsilon^{\mathrm{voigt}}_{j} - \epsilon^{\mathrm{voigt,dft}}_{j}
        \right)^2
    }
}} \nonumber \\
& + \chi_{\mathrm{rose}}\mathbf{L}^{\mathrm{Rose}} 
+ \chi_{\mathrm{elastic}}\mathbf{L}^{\mathrm{elastic}}
\end{align} 
where $N_{b}$ is the batch size and $N_i$ is the number of atoms in structure 
$i$. $\chi_{\mathrm{f}}$, $\chi_{\mathrm{s}}$, $\chi_{\mathrm{rose}}$ and 
$\chi_{\mathrm{elastic}}$ are overall weights of force, virial stress, rose EOS 
and elastic contributions. The superscript 'voigt' means that virial 
stress tensors should be converted to Voigt form. The Adam optimizer is used to 
minimize Equation \ref{eq:loss}. In most cases, we use 0.01 as the initial 
learning rate and the batch size ranges from 20 to 50. Since there are very few
adjustable parameters, the optimization typically needs very few epochs to 
converge. When the optimization is finished, the corresponding Lammps setfl 
potential file will be exported.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3. Results
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Results}
\label{sec:results}

The publicly Ni-Mo dataset is used to evaluate our approach. This dataset is 
provided by Shyue Ping Ong. It contains 3973 different Ni-Mo solids. All 
calculations were done by VASP with the PBE functional and the projector 
augmented-wave approach.

The optimized parameters are listed in the appendix.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Elementary Ni and Mo}
\label{sec:elementary}

We first use the two elementary datasets (Ni and Mo) to show our approach. These 
two datasets has 461 and 284 structures, respectively. 61 and 34 were randomly 
selected to form the test subsets. In these two experiments, 
$\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and $\chi_{\mathrm{elastic}}$ were 
set to 1, 3 and 0.05 respectively. The batch size was fixed to 25.

The original Zjw04 potentials were fitted to experimentally obtained quantities.
Thus, we use the relative MAE:
\begin{equation}
\mathbf{rMAE} = \frac{1}{N}\sum_{i=1}^{N}{
    | \frac{(E_i - E_i^{dft})}{N_i} - (E_{eq} - E_{eq}^{dft}) |
}
\end{equation}
to measure their performaces. Here $N_i$ is the number of atoms in structure 
$i$, $E_i$ is the Zjw04 energy of structure $i$ while $E_i^{dft}$ represents its 
corresponding DFT energy, $E_{eq}$ and $E_{eq}^{dft}$ refers to the equilibrium 
energy of Zjw04 and DFT, respectively.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: r(Mo-Mo)
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.64]{figures/r_MoMo_hist.png}
\caption{\label{fig:r_MoMo_hist} The energy-volume curves of the original EAM, 
SNAP, ML-EAM and DFT for the fcc Ni.}
\end{figure}

Figure \ref{fig:Mo_eam_adp} compares the $\rho(r)$, $F(\rho)$, $\phi(r)$, 
$\mu(r)$ and $\omega(r)$ functions. The differences between ML-EAM and ML-ADP 
$\rho(r)$, $\phi(r)$ and $F(\rho)$ are quite small. Both the angular-dependent 
dipole $\mu(r)$ and quadrupole $\omega(r)$ quickly descrease to zero at 2.5 \AA. 
However, the equilibrium spacing between two adjacent Mo atoms is 2.7 \AA. This 
may explain why ML-ADP does not outperform ML-EAM on this dataset: most 
$r^{\mathrm{Mo-Mo}}$ are beyond 2.5 \AA (Figure \ref{fig:r_MoMo_hist}).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.65]{figures/Ni_all_eos.png}
\caption{\label{fig:Ni_eos} The energy-volume curves of the original EAM, 
ML-EAM and DFT for the fcc Ni.}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.65]{figures/Mo_all_eos.png}
\caption{\label{fig:Mo_eos} The energy-volume curves of the original EAM, 
ML-EAM and DFT for the fcc Ni.}
\end{figure}

For the Ni dataset, the original Zjw04 already has a reasonable performace. 
The energy \textbf{rMAE} is 10.6 meV/atom and the force MAE is just 0.06 eV/\AA 
on the entire dataset. Our machine learning approach can further enhance the 
performace. ML-EAM achieves 4.1 meV/atom and 0.05 eV/\AA on the test set and 
3.9 meV/atom and 0.05 eV\AA on the entire dataset. ML-EAM is nearly as accurate 
as SNAP (1.2 meV/atom, 0.05 eV\AA), but later has around three orders of 
magnitude more computational expense. Figure \ref{fig:Ni_eam} demonstrates the 
$\rho(r)$, $F(\rho)$ and $\phi(r)$ functions before (Zjw04) and after machine 
learning optimization. The pairwise potential $\phi(r)$ almost remains the same 
while the embedding function $F(\rho)$ and the electron density function 
$\rho(r)$ changes significantly. Both $\rho(r)$ and $\phi(r)$ converge to zero 
at around $r=4.8$ \AA, which is rougly 2 times of the equilibrium spacing 
between nearest neighbors.

For the Mo dataset, the situation is quite different. The original Zjw04 has 
extremely large MAEs: the energy \textbf{rMAE} is $\sim 98$ meV/atom and the 
force MAE is more than 0.40 eV/\AA. Our method can significantly reduce the 
error: ML-EAM can achieve test MAEs of 21.9 meV/atom and 0.28 eV/\AA and overall
MAEs of 28.1 meV/atom and 0.32 eV/\AA. We also optimized an ADP potential for 
this dataset. Interestingly, ML-ADP achieves almost exactly the same MAEs with 
ML-EAM: 22.1 meV/atom and 0.27 eV\AA on the test set and 27.9 meV/atom and 0.33 
eV/\AA on the entire dataset. On the contrast, the SNAP method (13.2 meV/atom, 
0.25 eV/\AA) behaves better. But considering the computational cost, the 
optimized EAM and ADP are totally accepetable.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Mo - Material Properties
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}
\centering
\begin{tabular}{lcccccc}
\hline
                              & DFT  & SNAP          & EAM           & ML-EAM         & ML-ADP         & Exp. \\
\hline
$T_{m}$(K)                    &      & 3000          & 3750          & 2940           & 2880           & 2890 \\
$c_{11}$ (GPa)                & 472  & 473 (0.2\%)   & 457 (-3.2\%)  & 472 (0.0\%)    & 473 (0.2\%)    & 479  \\
$c_{12}$ (GPa)                & 158  & 152 (-3.8\%)  & 168 (6.3\%)   & 158 (0.0\%)    & 160 (1.2\%)    & 165  \\
$c_{44}$ (GPa)                & 106  & 107 (0.9\%)   & 116 (9.4\%)   & 106 (0.0\%)    & 105 (-0.9\%)   & 108  \\
$B_{\mathrm{VRH}}$ (GPa)      & 263  & 259 (-1.5\%)  & 264 (0.4\%)   & 262 (-0.4\%)   & 264 (0.4\%)    & 270  \\
$G_{\mathrm{VRH}}$ (GPa)      & 124  & 126 (1.6\%)   & 127 (2.4\%)   & 125 (0.8\%)    & 124 (0.0\%)    & 125  \\
$\mu$                         & 0.30 & 0.29 (-3.3\%) & 0.29 (-3.3\%) & 0.30 (0.9\%)   & 0.30 (0.0\%)   & 0.30 \\
$E_{v}$ (eV)                  & 2.87 & 2.61 (-9.1\%) & 3.02 (5.2\%)  & 1.16 (-20.5\%) & 1.71 (17.1\%)  &      \\
$E_{m}$ (eV)                  & 1.12 & 1.39 (24.1\%) & 1.54 (37.5\%) & 1.46 (30.5\%)  & 0.87 (-22.3\%) &      \\
$E_{a} = E_{v} + E_{m} $ (eV) & 3.99 & 4.00 (-0.1\%) & 4.56 (14.3\%) & 2.62 (1.6\%)   & 2.58 (0.0\%)   & 4.00 \\
\hline
\end{tabular}
\caption{\label{table:Mo_properties} Comparison of the calculated and 
experimental material properties of elementary bcc Mo. SNAP refers to the 
elementary Mo SNAP model.
}
\end{table*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Alloy EOS - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.5]{figures/alloy_eos.png}
\caption{\label{fig:alloy_eos} $\rho(r)$, $F(\rho)$, $\phi(r)$ of the original 
Zjw04 EAM, ML-EAM and ML-ADP. \textbf{(d)} shows the $\mu(r)$ and $\omega(r)$ of 
ML-ADP.}
\end{figure*}

Figure \ref{fig:Ni_eos} and \ref{fig:Mo_eos} plot the equation of state 
curves. The fitting region ranges from 84\% to 118\% of the corresponding 
equilibrium volume. For both crystals, the original EAM significantly 
underestimate the energy at both tensile and compress strains. However, our 
constrained machine learning approach can fix this problem. The 
energy-volume curves of ML-EAM and ML-ADP overalp with the DFT curves very well.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EAM/ADP - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[htp]
\centering
\includegraphics[scale=0.6]{figures/Mo_eam_adp.png}
\caption{\label{fig:Mo_eam_adp} $\rho(r)$, $F(\rho)$, $\phi(r)$ of the original 
Zjw04 EAM, ML-EAM and ML-ADP. \textbf{(d)} shows the $\mu(r)$ and $\omega(r)$ of 
ML-ADP.}
\end{figure}

Table \ref{table:Ni_properties} and \ref{table:Mo_properties} summerize the 
predicted material properties of fcc Ni and bcc Mo, including the melting point 
$T_{m}$, the elastic constants ($c_{11}$, $c_{12}$ and $c_{44}$), the vacancy 
formation energy $E_{v}$, the migration energy $E_{m}$ and the activation energy 
$E_{a}$. The original Zjw04 potentials were developed by fitting material 
properties so they already perform well on these metrics. Our ML-EAM and ML-ADP 
not only inherit this feature but also gain significant improvement on 
predicting elastic constants: the relative errors of the elastic constants as 
well as the Voigt-Reuss-Hill bulk modulus, shear modulus and homogeneous 
Poisson's ratio, for both bcc Mo and fcc Ni, are reduced from $\sim$10\% to 
0-2\%.

We further examined our ML-EAM and ML-ADP by calculating the melting temperature
of fcc Ni and bcc Mo using the solid-liquid coexistence approach. For fcc Ni, 
$30 \times 10 \times 10$ supercells were used and for bcc the supercells were 
$30 \times 15 \times 15$. The time step was set to 1 fs. The simulation for each 
temperature was carried out for at least 300 ps. For fcc Ni, $T_{m}$ remains the 
same for ML-EAM, 200 K lower than the experiment $T_m$. However, the calculated 
$T_{m}$ of Mo is greatly improved. The original Zjw04 $T_{m}$ is around 3750 K, 
800 K higher than experiment (2890 K). Both ML-EAM and ML-ADP can give very 
accurate melting temperature.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Mo-Ni}
\label{sec:alloy}

Secondly, we tested our method on the entire binary Mo-Ni dataset. 300 
structures were randomly splitted to form the test dataset and the rest 3673 
were used to train the potentials. $\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ 
and $\chi_{\mathrm{elastic}}$ were set to 1, 1 and 0.05 respectively. The batch 
size was increased to 50. Only the equilibrium fcc Ni and bcc Mo are included in 
$\Omega$ of Equation \ref{eq:cijkl_loss} and \ref{eq:rose_loss}.

The original EAM uses a simplified interpolation form to describe the pairwise 
interaction of A-B (Equation \ref{eq:zjw04_phi_ab}). However, we did a minor 
modification: by adding more fitting parameters we use Equation 
\ref{eq:zjw04_phi_aa} directly. Thus, in total, there are 44 learnable 
parameters for ML-EAM and 68 for ML-ADP. One should note that in this section, 
SNAP, ML-EAM and ML-ADP all refer to their corresponding binary form.

Table \ref{table:MAE} summarizes the energy and force MAEs of SNAP, ML-EAM and 
ML-ADP on the different subsets. Amazingly, ML-ADP becomes the overally best. 
Both the energy MAE (19.6 meV/atom) and the force MAE (0.20 eV/\AA) are smaller 
than those of the binary SNAP (22.5 meV/atom and 0.23 eV/\AA). 
The pure Mo is still a challenge for ML-ADP as the energy and force MAEs are 
noticeably large (still much smaller than the original Zjw04 EAM). The binary 
SNAP predicts well on pure Ni and Mo. But ML-ADP outperforms the SNAP model on 
binary phases (Ni$_3$Mo, Ni-doped Mo and especially the Mo-doped Ni).
On the contrast, ML-EAM performs significantly worse compared with ML-ADP. 
Interestingly, our results suggest that Mo-Ni dipole and quadrupole interactions 
are completely negligible (Figure S1 in the appendix). The Ni-Ni dipole and 
quadrupole interactions are an order of magnitude weaken than those of Mo-Mo.

Figure \ref{fig:alloy_eos} demonstrates the energy-volume curves of Ni$_3$Mo and
Ni$_4$Mo. The original Zjw04 EAM can not properly describe the energy-volume 
relations of these two crystals. The ML-ADP curves fit perfectly with the DFT 
curves. ML-EAM fits well with tensile strains but overestimates the energy under 
compress strains.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Mo-Ni MAE
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}
\centering
\begin{tabular}{lcccccccc}
\hline
                  & Model  & Mo   & Ni$_4$Mo & Ni$_3$Mo & Mo$_{\mathrm{Ni}}$ & Ni$_{\mathrm{Mo}}$ & Ni   & Overall \\
\hline
Energy (meV/atom) & SNAP   & 16.2 & 4.0      & 5.2      & 22.7               & 33.9               & 7.9  & 22.5    \\
                  & ML-EAM & 30.4 & 11.1     & 8.6      & 29.8               & 33.9               & 11.6 & 26.2    \\
                  & ML-ADP & 36.9 & 5.1      & 4.8      & 21.7               & 22.8               & 13.3 & 19.6    \\
\hline
Force (eV/\AA)    & SNAP   & 0.29 & 0.14     & 0.16     & 0.13               & 0.55               & 0.11 & 0.23    \\
                  & ML-EAM & 0.33 & 0.24     & 0.19     & 0.25               & 0.35               & 0.08 & 0.23    \\
                  & ML-ADP & 0.45 & 0.16     & 0.15     & 0.14               & 0.35               & 0.08 & 0.20    \\
\hline
\end{tabular}
\caption{\label{table:MAE}
Comparion of the MAEs in predicted energies (mev/atom) and forces (eV/\AA) 
relative to the DFT on the entire Mo-Ni dataset.}
\end{table*}

Table \ref{table:NiMo_elastic_constants} summarizes the elastic constants 
predictions of Ni$_3$Mo and Ni$_4$Mo. One should note that these two crystals 
were not included in $\Omega$ of Equation \ref{eq:cijkl_loss} and 
\ref{eq:rose_loss} but their elastic properties were used to tune the binary 
SNAP model. The original Zjw04 EAM performs obviously poorly on predicting 
elastic constants. Some absolute percentage errors can even exceed 100\%. With 
our machine learning correction, ML-EAM, especially ML-ADP, can give much more  
accurate predictions. For the Ni$_3$Mo crystal, the average absolute percentage 
error is reduced from 40\% (Zjw04) to 20\% (ML-ADP). ML-ADP even beats SNAP on 4 
($c_{11}$, $c_{12}$, $c_{22}$ and $c_{3}$) of the 7 elastic metrics. For the 
Ni$_4$Mo crystal, the average absolute percentage error is also descreased.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Mo-Ni Elastic Constants
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table*}
\centering
\begin{tabular}{lccccc}
\hline
                   & DFT  & SNAP          & EAM           & ML-EAM        & ML-ADP        \\
\hline
Ni$_3$Mo \\
$c_{11}$           & 385  & 420 (9.1\%)   & 195 (-49.4\%) & 403 (4.7\%)   & 374 (-2.9\%)  \\
$c_{12}$           & 166  & 197 (18.7\%)  & 98 (-41.0\%)  & 208 (25.3\%)  & 160 (-3.6\%)  \\
$c_{13}$           & 145  & 162 (11.7\%)  & 98 (-32.4\%)  & 230 (58.6\%)  & 176 (21.4\%)  \\
$c_{22}$           & 402  & 360 (10.7\%)  & 351 (-12.7\%) & 443 (10.2\%)  & 386 (-4.0\%)  \\
$c_{23}$           & 131  & 145 (-10.4\%) & 107 (-18.3\%) & 272 (107.6\%) & 214 (63.3\%)  \\
$c_{33}$           & 402  & 408 (1.5\%)   & 295 (-26.6\%) & 474 (17.9\%)  & 398 (-0.5\%)  \\
$c_{44}$           & 94   & 84 (-10.4\%)  & 36 (-61.7\%)  & 25 (-73.4\%)  & 52 (-44.6\%)  \\
$B_{\mathrm{VRH}}$ & 230  & 243 (5.7\%)   & 156 (-32.2\%) & 302 (31.3\%)  & 250 (8.7\%)   \\
$G_{\mathrm{VRH}}$ & 89   & 100 (12.4\%)  & 61 (-31.5\%)  & 57 (36.0\%)   & 66 (25.8\%)   \\
$\mu$              & 0.33 & 0.32 (-3.0\%) & 0.33 (0.0\%)  & 0.41 (24.2\%) & 0.38 (15.1\%) \\
\hline
Ni$_4$Mo \\
$c_{11}$           & 300  & 283 (-5.7\%)  & 172 (-42.7\%) & 282 (-6.0\%)  & 342 (14.0\%)  \\
$c_{12}$           & 186  & 179 (-3.8\%)  & 158 (-15.1\%) & 125 (-32.8\%) & 217 (16.7\%)  \\
$c_{22}$           & 313  & 326 (4.2\%)   & 158 (-49.5\%) & 282 (-10.0\%) & 342 (9.3\%)   \\
$c_{23}$           & 166  & 164 (-1.2\%)  & 80 (-51.8\%)  & 156 (-6.0\%)  & 232 (39.8\%)  \\
$c_{44}$           & 130  & 126 (-3.1\%)  & 125 (-3.8\%)  & 136 (4.6\%)   & 117 (-10.0\%) \\
$B_{\mathrm{VRH}}$ & 223  & 220 (-1.3\%)  & 161 (-27.8\%) & 178 (-20.1\%) & 257 (15.2\%)  \\
$G_{\mathrm{VRH}}$ & 91   & 95 (4.4\%)    & -156 (-162\%) & 80 (-12.1\%)  & 79 (-13.2\%)  \\
$\mu$              & 0.33 & 0.31 (-6.1\%) & 0.70 (112\%)  & 0.31 (-6.1\%) & 0.36 (9.1\%)  \\
\hline
\end{tabular}
\caption{\label{table:NiMo_elastic_constants}
Comparion of elastic constants ($c_{ij}$, GPa), Voigt-Reuss-Hill bulk modulus 
($B_{\mathrm{VRH}}$, GPa), Voigt-Reuss-Hill shear modulus ($G_{\mathrm{VRH}}$, 
GPa) and homogeneous Poisson's ratio ($\mu$) for fcc Ni, fcc Mo and binary 
alloys Ni$_{3}$Mo and Ni$_{4}$Mo.
}
\end{table*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4. Discussions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Discussions}
\label{sec:discussions}

The above results suggest that with machine learning and essential physical 
constraints, for the fcc and bcc solids, the traditional empirical potentials, 
EAM/ADP, can be improved to be almost as accurate as the SNAP method. But, 
EAM/ADP are around $10^2$ to $10^3$ times faster than SNAP. To understand this,
we can start from the many-body expansion (MBE) scheme.

MBE is one of the most widely used schemes for fitting potential energy 
surfaces. In the many-body expansion scheme, the total energy of a system with N 
atoms can be expressed as the sum of all k-body terms where $k \le 1$:
\begin{equation}
\label{eq:many_body_expansion}
E = 
\sum_{i}^{\mathrm{C^N_1}}{E^{(1)}_{i}} +
\sum_{i,j}^{\mathrm{C^N_2}}{E^{(2)}_{ij}} + 
\sum_{i,j,k}^{\mathrm{C^N_3}}{E^{(3)}_{ijk}} + \cdots 
\end{equation}
where $\mathrm{C^N_k}$ is the binomial coefficient and $E^{(k)}$ represents the 
k-body contribution. In many cases, higher-order terms like $E_{ij}$ or 
$E_{ijk}$ are symmetric: $E_{ij}=E_{ji}$, $E_{ijk}=E_{ikj}=E_{jik}=\cdots$. 
Thus, Equation \ref{eq:many_body_expansion} can be further transformed to:
\begin{align}
\label{eq:MBE_atomic}
E & = \sum_{i}^{\mathrm{N}}{\left(
    E^{(1)}_{i} + 
    \frac{1}{2!}\sum_{j \ne i}^{\mathrm{N}}{E^{(2)}_{ij}} +
    \frac{1}{3!}\sum_{j \ne i}^{\mathrm{N}}{
        \sum_{k \ne i,j}^{\mathrm{N}}{E^{(3)}_{ijk}}
    } +
    \cdots
\right)} \nonumber \\
& = \sum_{i}^{\mathrm{N}}{E_{i}}
\end{align}
where $E_{i}$ is the atomic energy of atom $i$.

In fact, both the EAM and the SNAP formalisms can be considered as variants of 
MBE truncated at $k=2$. For the EAM method, the embedding contribution $F(\rho)$ 
serves as the one-body term and the overall pairwise contribution 
$\frac{1}{2}\sum_{j\ne i}{\phi(r_{ij})}$ is just a plain transcription of the 
two-body term in Equation \ref{eq:MBE_atomic}. The ADP formalism, which 
introduces extra pairwise dipole and quadrupole interactions, does not change
this nature. For the SNAP method, the atomic energy is calculated with the 
following equation: 
\begin{equation}
\label{eq:snap_formalism}
E_i = \beta_{0}^{\alpha_{i}} + \sum_{k}{\beta_{k}^{\alpha_{i}}B_{k}^i}
\end{equation} 
where $\beta_{0}^{\alpha_i}$ and $\beta_{k}^{\alpha_i}$ are learnable scalar
parameters, $\alpha_i$ indicates the element type of atom $i$. $B_{k}^{i}$ is a 
bispectrum coefficient. The calculation of $B_{k}^{i}$ is very complicated but 
it is still a two-body term. The overall two-body contribution is a linear 
combination of $B_{k}^{i}$. Thus, the SNAP formalism can be viewed as an 
assemble of a simplified one-body term with a complicated two-body term. On the 
other hand, the EAM formalism is built up with a 'complicated' one-body term and 
a rather simple two-body term. Hence, for the fcc and bcc systems, both methods 
perform similarly.

As we analysed in Section \ref{sec:transformation}, the fundamental atomic 
descriptor of EAM, the $N \times N^{\mathrm{nl}} \times 4$ matrix $\mathbf{G}$, 
is composed of just raw interatomic distances. However, interatomic distances
alone may not uniquely describe a structure. To avoid this problem, the cutoff 
was raised up to 6.5 \AA. As a comparison, the overall $r_{cut}$ is 4.6 \AA for 
the binary SNAP. But such descriptor is still too rough to describe complicated 
potential energy surfaces (e.g. bcc Mo). To further enhance EAM/ADP, perhaps we 
shall consider adding real three-body terms (e.g. Tersoff).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 5. Conclusions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Conclusions}
\label{sec:conclusions}

To conclude, we have successfully combined empirical potentials (EAM/ADP) with 
machine learning. The machine learning approaches (big data, SGD, etc), together 
with physical constraints (Rose EOS, elastic constants) can significantly boost
the performaces of EAM/ADP. For the fcc Ni, bcc Mo and Mo-Ni alloys, ML-EAM and
ML-ADP can be as accurate as SNAP, while later method is orders of magnitude 
slower. Our work also indicates a new route to design and develop machine 
learning interatomic potentials \textemdash the machine learning enhanced 
empirical potential approach.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Acknowlegements
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section*{Acknowledgments}
\label{sec:acknowledgments}

This work was supported by the National Key Research and Development Program of 
China under Grant No. 2016YFB0201204, the Science Challenge Project under Grant 
No. TZ2018002 and the National Natural Science Foundation of China under Grant 
No. U1630250.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% References
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\bibliography{manuscript.bib}

\end{document}